---
title: "Phytoplankton Indicators"
author: "Brandon Beltz"
date: "`r Sys.Date()`"
output:
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ecodata)
library(ecodata25)
library(dplyr)
library(knitr)

# Replace 'chl_pp' with intended dataset
dataset<-ecodata::chl_pp
dataset24<-ecodata25::chl_pp

# Define dataset-specific arguments for plot functions
report<-list("MidAtlantic","NewEngland")
variable<-list("chl","pp","size")
plottype<-list("weekly","monthly")
epu<-list("MAB","GOM","GB")
```

# Phytoplankton Figures {.tabset}

## `r paste(report[1],variable[1],plottype[1],epu[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[1], n = 10, varName = variable[1], plottype = plottype[1], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[1], n = 10, varName = variable[1], plottype = plottype[1], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[1],plottype[1],epu[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[1], plottype = plottype[1], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[1], plottype = plottype[1], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[1],plottype[1],epu[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[1], plottype = plottype[1], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[1], plottype = plottype[1], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[1],plottype[2],epu[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[1], n = 10, varName = variable[1], plottype = plottype[2], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[1], n = 10, varName = variable[1], plottype = plottype[2], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[1],plottype[2],epu[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[1], plottype = plottype[2], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[1], plottype = plottype[2], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[1],plottype[2],epu[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[1], plottype = plottype[2], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[1], plottype = plottype[2], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[2],plottype[1],epu[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[1], n = 10, varName = variable[2], plottype = plottype[1], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[1], n = 10, varName = variable[2], plottype = plottype[1], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[2],plottype[1],epu[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[2], plottype = plottype[1], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[2], plottype = plottype[1], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[2],plottype[1],epu[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[2], plottype = plottype[1], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[2], plottype = plottype[1], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[2],plottype[2],epu[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[1], n = 10, varName = variable[2], plottype = plottype[2], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[1], n = 10, varName = variable[2], plottype = plottype[2], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[2],plottype[2],epu[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[2], plottype = plottype[2], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[2], plottype = plottype[2], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[2],plottype[2],epu[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[2], plottype = plottype[2], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[2], plottype = plottype[2], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[3],plottype[1],epu[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[1], n = 10, varName = variable[3], plottype = plottype[1], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[1], n = 10, varName = variable[3], plottype = plottype[1], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[3],plottype[1],epu[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[3], plottype = plottype[1], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[3], plottype = plottype[1], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[3],plottype[1],epu[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[3], plottype = plottype[1], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[3], plottype = plottype[1], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[3],plottype[2],epu[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[1], n = 10, varName = variable[3], plottype = plottype[2], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[1], n = 10, varName = variable[3], plottype = plottype[2], EPU = epu[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[3],plottype[2],epu[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[3], plottype = plottype[2], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[3], plottype = plottype[2], EPU = epu[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[3],plottype[2],epu[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_chl_pp(report = report[2], n = 10, varName = variable[3], plottype = plottype[2], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_chl_pp(report = report[2], n = 10, varName = variable[3], plottype = plottype[2], EPU = epu[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```
