---
title: "Engagement Indices"
author: "Brandon Beltz"
date: "`r Sys.Date()`"
output:
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ecodata)
library(ecodata25)
library(dplyr)
library(knitr)

# Replace 'engagement' with intended dataset
dataset<-ecodata::engagement
dataset24<-ecodata25::engagement

# Define dataset-specific arguments for plot functions
report<-list("MidAtlantic","NewEngland")
variable<-list("Commercial", "Recreational")
plottype<-list("plot","Social", "Economic", "Gentrification")
#epu<-list()
```

# Dataset Metrics {.tabset}
```{r echo=FALSE, include=TRUE}
datastats<-data.frame(Metric=c("First Year","Final Year","Series Length","Minimum Value","Maximum Value","Mean Value","Median Value","# NAs"),
                      Last_Year=c(min(dataset24$Time),max(dataset24$Time),max(dataset24$Time)-min(dataset24$Time),min(dataset24$Value, na.rm = T),max(dataset24$Value, na.rm = T),mean(dataset24$Value, na.rm = T),median(dataset24$Value, na.rm = T),sum(is.na(dataset24$Value))),
                      Current_Year=c(min(dataset$Time),max(dataset$Time),max(dataset$Time)-min(dataset$Time),min(dataset$Value, na.rm = T),max(dataset$Value, na.rm = T),mean(dataset$Value, na.rm = T),median(dataset$Value, na.rm = T),sum(is.na(dataset24$Value))))
datastats$Difference<-datastats$Current_Year-datastats$Last_Year

kable(datastats, digits = 0)
```

## `r paste(report[1],variable[1],plottype[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[1], varName = variable[1], plottype = plottype[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[1], varName = variable[1], plottype = plottype[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[1],plottype[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[2], varName = variable[1], plottype = plottype[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[2], varName = variable[1], plottype = plottype[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[1],plottype[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[1], varName = variable[1], plottype = plottype[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[1], varName = variable[1], plottype = plottype[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[1],plottype[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[2], varName = variable[1], plottype = plottype[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[2], varName = variable[1], plottype = plottype[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[1],plottype[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[1], varName = variable[1], plottype = plottype[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[1], varName = variable[1], plottype = plottype[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[1],plottype[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[2], varName = variable[1], plottype = plottype[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[2], varName = variable[1], plottype = plottype[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[1],plottype[4])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[1], varName = variable[1], plottype = plottype[4]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[1], varName = variable[1], plottype = plottype[4]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[1],plottype[4])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[2], varName = variable[1], plottype = plottype[4]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[2], varName = variable[1], plottype = plottype[4]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[2],plottype[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[1], varName = variable[2], plottype = plottype[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[1], varName = variable[2], plottype = plottype[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[2],plottype[1])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[2], varName = variable[2], plottype = plottype[1]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[2], varName = variable[2], plottype = plottype[1]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[2],plottype[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[1], varName = variable[2], plottype = plottype[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[1], varName = variable[2], plottype = plottype[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[2],plottype[2])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[2], varName = variable[2], plottype = plottype[2]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[2], varName = variable[2], plottype = plottype[2]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[2],plottype[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[1], varName = variable[2], plottype = plottype[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[1], varName = variable[2], plottype = plottype[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[2],plottype[3])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[2], varName = variable[2], plottype = plottype[3]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[2], varName = variable[2], plottype = plottype[3]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[1],variable[2],plottype[4])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[1], varName = variable[2], plottype = plottype[4]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[1], varName = variable[2], plottype = plottype[4]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```

## `r paste(report[2],variable[2],plottype[4])`
```{r, fig.show="hold", out.width="50%", echo=FALSE, message=FALSE, warning=FALSE}
ecodata::plot_engagement(report = report[2], varName = variable[2], plottype = plottype[4]) + 
  ggplot2::ggtitle("SOE 2026") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))

ecodata25::plot_engagement(report = report[2], varName = variable[2], plottype = plottype[4]) + 
  ggplot2::ggtitle("SOE 2025") + ggplot2::theme(plot.title = ggplot2::element_text(size = 20))
```
