---
title: "comdat exploration"
format: html
# format: pdf
---

## Exploring who manages the things that are caught

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

### Playing with plots

Create the data with the right resolution:
```{r}
new_comdat <- ecodata::comdat |>
  dplyr::mutate(
    category = dplyr::case_when(
      stringr::str_detect(Var, "Landings") ~ "Landings",
      stringr::str_detect(Var, "Revenue") ~ "Revenue",
      TRUE ~ "Other"
    ),
    details = dplyr::case_when(
      stringr::str_detect(Var, "US only") ~ "US only",
      stringr::str_detect(Var, "Seafood Landings") ~ "Seafood Landings",
      stringr::str_detect(Var, "Landings weight$") ~ "Landings weight",
      stringr::str_detect(Var, "Revenue") ~ "Revenue",
      TRUE ~ NA_character_
    ),
    managed_by = dplyr::case_when(
      stringr::str_detect(Var, "NEFMC") ~ "NEFMC",
      stringr::str_detect(Var, "MAFMC") ~ "MAFMC",
      stringr::str_detect(Var, "JOINT") ~ "JOINT",
      # "Other managed" is the species not managed by NE or Mid
      # https://github.com/NEFSC/READ_EDAB_SOE_Workflows/blob/main/R/comdat.R#L206
      # could be state managed or unmanaged
      stringr::str_detect(Var, "Other managed") ~ "Not federally managed",
      TRUE ~ NA_character_
    ),
    guild = dplyr::case_when(
      stringr::str_detect(Var, "Apex Predator") ~ "Apex Predator",
      stringr::str_detect(Var, "Benthivore") ~ "Benthivore",
      stringr::str_detect(Var, "Benthos") ~ "Benthos",
      stringr::str_detect(Var, "Other Other") ~ "Other",
      stringr::str_detect(Var, "Piscivore") ~ "Piscivore",
      stringr::str_detect(Var, "Planktivore") ~ "Planktivore",
      TRUE ~ "All"
    )
  ) |>
  dplyr::filter(
    !EPU %in% c("SS", "OTHER", "Other"),
    Time > 1981,
    !is.na(managed_by),
    guild != "All"
  )
```

#### Landings and revenue by jurisdiction x EPU

Plot by council and EPU:
```{r}
## plot landings by council x EPU ----

comdat_council_epu <- new_comdat |>
  dplyr::filter(
    !is.na(managed_by),
    details == "Landings weight" |
      details == "Revenue"
  ) |>
  dplyr::group_by(Time, EPU, managed_by, category) |>
  dplyr::summarise(total_value = sum(Value, na.rm = TRUE)) |>
  dplyr::ungroup()

### line plot ----
comdat_council_epu |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = total_value, color = managed_by)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(category),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)

### stacked bar plot ----
comdat_council_epu |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = total_value, fill = managed_by)) +
  ggplot2::geom_col() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(category),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)

### proportions ----

prop_all_epu <- comdat_council_epu |>
  dplyr::group_by(Time, EPU, category) |>
  dplyr::mutate(proportion = total_value / sum(total_value, na.rm = TRUE)) |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = proportion, fill = managed_by)) +
  ggplot2::geom_col() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(category),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)
```

#### Seafood and non-seafood landings by jurisdiction x EPU

```{r}
# find non-seafood landings ----
non_seafood_landings <- new_comdat |>
  dplyr::filter(
    !is.na(managed_by),
    details == "Landings weight" |
      details == "Seafood Landings"
  ) |>
  dplyr::select(-Var) |>
  tidyr::pivot_wider(names_from = details, values_from = Value) |>
  dplyr::mutate(
    Value = `Landings weight` - `Seafood Landings`,
    details = "Non-Seafood Landings"
  ) |>
  dplyr::select(-`Landings weight`, -`Seafood Landings`)

seafood_dat <- new_comdat |>
  dplyr::filter(
    !is.na(managed_by),
    details == "Seafood Landings"
  ) |>
  dplyr::bind_rows(non_seafood_landings) |>
  dplyr::filter(
    !is.na(managed_by),
    details == "Seafood Landings" |
      details == "Non-Seafood Landings"
  ) |>
  dplyr::group_by(Time, EPU, managed_by, details) |>
  dplyr::summarise(total_value = sum(Value, na.rm = TRUE)) |>
  dplyr::ungroup()

## plot landings of seafood and non-seafood ----

### line plot ----
seafood_dat |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = total_value, color = managed_by)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(details)
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)


### stacked bar plot ----
seafood_dat |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = total_value, fill = managed_by)) +
  ggplot2::geom_col() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(details),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)


### proportions ----

seafood_dat |>
  dplyr::group_by(Time, EPU, details) |>
  dplyr::mutate(proportion = total_value / sum(total_value, na.rm = TRUE)) |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = proportion, fill = managed_by)) +
  ggplot2::geom_col() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(details),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)
```

#### Total landings and revenue by jurisdiction

```{r}
new_dat <- comdat_council_epu |>
  dplyr::bind_rows(seafood_dat |>
                     dplyr::filter(details == "Seafood Landings") |>
                     dplyr::mutate(category = "Landings")) |>
  dplyr::mutate(details = dplyr::case_when(category == "Landings" & is.na(details) ~ "Total Landings",
                                           category == "Revenue" ~ "Revenue",
                                           TRUE ~ details))


new_dat |>
  dplyr::group_by(Time, managed_by, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = details)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(managed_by),
    rows = ggplot2::vars(category),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)
```

#### Total landings and revenue by EPU

```{r}
new_dat |>
  dplyr::group_by(Time, EPU, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = details)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(category),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)
```


### Plots for SOE

#### Magnitude of landings and revenue by jurisdiction 

Mid example:
```{r}
new_dat |>
  dplyr::filter(managed_by == "JOINT" | managed_by == "MAFMC",
                category == "Landings") |>
  dplyr::group_by(Time, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = details)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Landings of MAFMC-managed species")
```

```{r}
new_dat |>
  dplyr::filter(managed_by == "JOINT" | managed_by == "MAFMC",
                category == "Revenue") |>
  dplyr::group_by(Time, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Revenue of MAFMC-managed species")
```

NE example:
```{r}
new_dat |>
  dplyr::filter(managed_by == "JOINT" | managed_by == "NEFMC",
                category == "Revenue") |>
  dplyr::group_by(Time, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = details)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Landings and Revenue of NEFMC-managed species")
```

#### Magnitude of landings and revenue by EPU

Again, just showing Mid as an example:

```{r}
new_dat |>
  dplyr::filter(EPU == "MAB",
                category == "Landings") |>
  dplyr::group_by(Time, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = details)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Landings in the Mid-Atlantic Bight")

new_dat |>
  dplyr::filter(EPU == "MAB",
                category == "Revenue") |>
  dplyr::group_by(Time, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Revenue in the Mid-Atlantic Bight")
```

Georges Bank:

```{r}
new_dat |>
  dplyr::filter(EPU == "GB") |>
  dplyr::group_by(Time, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = details)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_grid(
    rows = ggplot2::vars(category),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Total landings and revenue in Georges Bank")
```

Gulf of Maine:

```{r}
new_dat |>
  dplyr::filter(EPU == "GOM") |>
  dplyr::group_by(Time, category, details) |>
  dplyr::summarise(total = sum(total_value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = details)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_grid(
    rows = ggplot2::vars(category),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Total landings and revenue in the Gulf of Maine")
```

#### Proportion of landings and revenue by jurisdiction x EPU

```{r}
comdat_council_epu |>
  dplyr::filter(category == "Landings") |>
  dplyr::group_by(Time, EPU, category) |>
  dplyr::mutate(proportion = total_value / sum(total_value, na.rm = TRUE)) |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = proportion, fill = managed_by)) +
  ggplot2::geom_col() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Proportion of landings by council jurisdiction and EPU")

comdat_council_epu |>
  dplyr::filter(category == "Revenue") |>
  dplyr::group_by(Time, EPU, category) |>
  dplyr::mutate(proportion = total_value / sum(total_value, na.rm = TRUE)) |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = proportion, fill = managed_by)) +
  ggplot2::geom_col() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Proportion of revenue by council jurisdiction and EPU")
```

### Guild plots

```{r}

guild_dat <- new_comdat |> 
  dplyr::filter(
    !is.na(managed_by),
    details == "Landings weight" |
      details == "Revenue"
  ) |>
  dplyr::filter(category == "Landings",
                guild != "All")


guild_dat |>
  dplyr::group_by(Time, EPU, guild, managed_by) |>
  dplyr::summarise(total_value = sum(Value, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::group_by(Time, EPU, managed_by) |>
  dplyr::mutate(proportion = total_value / sum(total_value, na.rm = TRUE)) |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = proportion, fill = guild)) +
  ggplot2::geom_col() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(managed_by),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)
```

```{r}
guild_dat |>
  dplyr::group_by(Time, EPU, guild, managed_by) |>
  dplyr::summarise(total_value = sum(Value, na.rm = TRUE)) |>
  ggplot2::ggplot(ggplot2::aes(x = Time, y = total_value, fill = guild)) +
  ggplot2::geom_col() +
  ggplot2::facet_grid(
    cols = ggplot2::vars(EPU),
    rows = ggplot2::vars(managed_by),
    scales = "free_y"
  ) +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma)
```

Landings of MAFMC species by guild

```{r}
guild_dat |>
  dplyr::filter(managed_by == "JOINT" | managed_by == "MAFMC",
                category == "Landings") |>
  dplyr::group_by(Time, category, guild) |>
  dplyr::summarise(total = sum(Value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = guild)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Landings of MAFMC-managed species by guild")
```

Landings in MAB by guild

```{r}
guild_dat |>
  dplyr::filter(EPU == "MAB",
                category == "Landings") |>
  dplyr::group_by(Time, category, guild) |>
  dplyr::summarise(total = sum(Value, na.rm = TRUE))  |>
  ggplot2::ggplot(ggplot2::aes(x = Time,
                               y = total,
                               color = guild)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::theme_bw() +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::ggtitle("Landings in the MAB by guild")
```
