---
title: 'State of the Ecosystem: Indicator Processing'
author: Sean Hardison
output_format: html_document
---

## Introduction
The purpose of this file is to document State of the Ecosystem (SOE) indicator data processing. To update existing data sets, set the `save_clean` parameter in the Rmarkdown set-up chunk to `TRUE`. Raw data for these indicators are available in the file directory `soe/inst/extdata`, and libraries required to process indicator data sets are shown below.

```{r, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align='center')

#Required libraries
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(here)
library(zoo)
library(kableExtra)

raw.dir <- here("data-raw","inst","extdata") #raw data directory
clean.dir <- here("data") #output directory for cleaned data

save_clean <- TRUE

```


## Data sets

### Surface winds

These data are sourced from the [NCEP North American Regional Reanalysis (NARR)](https://www.esrl.noaa.gov/psd/data/gridded/data.narr.monolevel.html), extending from January 1979 to September 2018. 

```{r surface_winds_vars, echo = FALSE}

fname <- "NCEP NARR surface wind; TKE; HLCY, monthly, 1979-2018, V1.csv"

vars <- data.frame(Variable = c("Wind speed",
                                "Wind direction",
                                "Turbulent kinetic energy",
                                "Storm relative helicity"),
                   Name = c("uwnd",
                            "vwnd",
                            "tke",
                            "hlcy"),
                   Units = c("m sec^-1^",
                             "°",
                             "J kg^-1^",
                             "m^2^sec^-2^"))

kable(vars, caption = paste0('Variables in "',fname,'"'))%>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center") %>%
  column_spec(3, width = "5cm")
```


Variables included in these data are surface wind speed and direction (*uwnd* and *vwnd* respectively), surface turbulent kinetic energy (TKE), and storm relative helicity (HLCY). An indicator for total wind speed is calculated below as
$$\textrm{TWS} = \sqrt{u^2 + v^2}$$.

**Filename**: NCEP NARR surface wind; TKE; HLCY, monthly, 1979-2018, V1.csv  
**Contributor**: Vincent Saba, (vincent.saba@noaa.gov)

```{r surface_winds}
# Read in raw data
d <- read.csv(file.path(raw.dir,"NCEP NARR surface wind; TKE; HLCY, monthly, 1979-2018, V1.csv"))

# Processing steps for all wind speed data
wind_clean1 <- d  %>% gather(., Var, Value, GB.uwnd:MAB.tke) %>% #convert wide to long
  dplyr::rename(Time = Month.Year) %>% #rename time variable
  separate(Var, c("EPU","Var"),"\\.") %>% #separate out EPU from variable names
  mutate(Time = as.yearmon(as.Date(.$Time,Time, format = "%d-%b-%y"), "%m/%Y"), #Convert to date format
         Units = plyr::mapvalues(Var, from = unique(Var), to = c(rep("J/kg",2),"m^2/sec^2","J/kg")), #add units
         Time, season = plyr::mapvalues(month(Time), from = seq(1,12,1), #Get season
                                         to = c(rep("winter",3),
                                                rep("spring",3),
                                                rep("summer",3),
                                                rep("fall",3)))) 

# Calculate total wind speed from u and v components
total_wind_speed <- wind_clean1 %>% 
  filter(Var == "uwnd"| Var == "vwnd") %>% #select variables
  spread(., Var, Value) %>% #convert to wide for calculating tws
  mutate(`total wind speed` = sqrt(uwnd^2 + vwnd^2)) %>%  #tws
  select(-uwnd, -vwnd) %>% #start processing back to SOE format
  gather(.,Var, Value, `total wind speed`) #convert to long

wind_clean <- rbind(wind_clean1, total_wind_speed)
wind_clean <- wind_clean %>% unite(., Var, c(Var, season), sep = " ") #merge season into Var column

if (save_clean){
save(wind_clean, file =
       file.path(clean.dir, "wind_clean.Rdata"))
}

```

### Slopewater proportions

Slopewater proportions give the percent total of water type observed in the deep Northeast Channel (150-200 m depth). 

```{r slopewater_prop_vars, echo = FALSE}

fname <- "slopewater_proportions.csv"

vars <- data.frame(Variable = c("Warm Slope Water proportion",
                                "Labrador Shelf Slope Water proportion"),
                   Names = c("WSW",
                             "LSLW"),
                   Units = c("unitless",
                             "unitless"))

kable(vars, caption = paste0('Variables in "',fname,'"'))%>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center") %>%
  column_spec(2, width = "5cm")
```

Raw data fields correspond to year, water mass flavor (WSW = Warm Slope Water, LSLW = Labrador Slope Water), and proportion of total expressed as a percentage. 

**Filename**: slopewater_proportions.csv  
**Contributor**: Paula Fratantoni (paula.fratantoni@noaa.gov)

```{r slopewater_proportions}
d <- read.csv(file.path(raw.dir,"slopewater_proportions.csv"))

slopewater <- d %>%
  dplyr::rename(Time = year, Var = water.mass.flavor, Value = prop) %>% 
  mutate(EPU = "all", Units = "unitless", Var2 = "proportion ne channel") %>% 
  unite(.,Var,c(Var,Var2), sep = " ")
  
if (save_clean){
save(slopewater, file =
       file.path(clean.dir, "slopewater_proportions.Rdata"))
}

```
 
### Ocean temperature (*in situ*)

These data include *in situ* regional time series of both surface and bottom water temperatures on the Northeast Shelf. Raw data is split into four files by EPU (SS, GOM, GB, and MAB).

```{r ocean_temp_insitu_vars, echo = FALSE}

fname <- "Eco{EPU}_core_Ttopbot.csv"

vars <- data.frame(Variable = c("SST anomaly",
                                "Reference SST (1981-2010)",
                                "Bottom temp. anomaly",
                                "Reference BT (1981-2010)"),
                   Names = c("Tsfc_anom",
                             "Tsfc_ref",
                             "Tbot_anom",
                             "Tbot_ref"),
                   Units = c("°C",
                             "°C",
                             "°C",
                             "°C"))

kable(vars, caption = paste0('Variables in "',fname,'"'))%>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center") %>%
  column_spec(2, width = "5cm")
```

**Filenames**: EcoSS_core_Ttopbot.csv, EcoGoM_core_Ttopbot.csv, EcoGB_core_Ttopbot.csv, EcoMAB_core_Ttopbot.csv  
**Contributor**: Paula Fratantoni (paula.fratantoni@noaa.gov)

```{r ocean_temperature}
ss <- read.csv(file.path(raw.dir,"EcoSS_core_Ttopbot.csv")) %>% mutate(EPU = "SS")
gom <- read.csv(file.path(raw.dir,"EcoGoM_core_Ttopbot.csv")) %>% mutate(EPU = "GOM")
gb <- read.csv(file.path(raw.dir,"EcoGB_core_Ttopbot.csv")) %>% mutate(EPU = "GB")
mab <- read.csv(file.path(raw.dir,"EcoMAB_core_Ttopbot.csv")) %>% mutate(EPU = "MAB")

oce <- rbind(ss, gom, gb, mab) %>% #bind all
  dplyr::rename(Time = decimal.year, Var = variable.name, Value = temperature) %>% #rename
  mutate(Units = "degreesC", Time = format(date_decimal(Time), "%Y-%m-%d")) #add units


```



