---
title: 'State of the Ecosystem: Macrofauna Indicator Processing & Visualization'
author: Ecosystems Dynamics and Assessment Branch
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Required libraries
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(here)
library(zoo)
library(kableExtra)
library(sf)
library(rgdal)
library(raster)
library(sp)
library(gridExtra)
library(htmlwidgets)
library(tseries)
library(plotly)

#Data directories
raw.dir <- here("inst","extdata") #raw data directory
clean.dir <- here("data") #output directory for cleaned data
gis.dir <- here("inst","extdata","gis")
sample.dir <- here("inst","extdata","sample")

#CRS
crs <- "+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"

#Write output to file
save_clean <- F

#Execute spatial processing (must be TRUE to write clean data to file). If FALSE, will load sample data from file for plotting example
spatial_processing <- F

```

```{css, echo = FALSE}
pre code, pre, code {
  white-space: pre !important;
  overflow-y: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
  max-height: 400px;
}
```

## Introduction
The purpose of this report is to document State of the Ecosystem (SOE) **human dimensions** indicator data processing. All R code used to process and visualize the following data sets is self-contained in the [Rmarkdown document](https://github.com/NOAA-EDAB/soe/blob/master/inst/Rmd/process_humandimensions.Rmd) associated with this HTML file. To run and update data sets in this document, set the `save_clean` parameter in the set-up chunk to `TRUE`. Raw data for these indicators are available in the file directory `soe/inst/extdata`.

## Data sets

### Mid-Atlantic Aquaculture {.tabset .tabset-fade}

New aquaculture data for the 2018-2019 Mid-Atlantic SOE were added directly from newly published reports. Here I download the existing data from the NEFSC ERDDAP server and add the newest values.

```{r aquaculture-processing}

#Download and process-----------------------------------------------------
aqua_url <- "https://comet.nefsc.noaa.gov/erddap/tabledap/aquaculture_soe_v1.csv?Time,Value,Units,Var,EPU"

d <- sprintf(aqua_url) %>% 
  purrr::map(function(x) {readr::read_csv(url(x))}) %>% 
  as.data.frame() %>% 
  filter(!is.na(Value))

va <- d %>%
  filter(Var == "aquaculture VA oyster harvest N") %>% 
  add_row(., Time = 2017,
          Value = 38900000,
          Var = "aquaculture VA oyster harvest N",
          EPU = "MAB",
          Units = "N")

nj <- d %>%
  filter(Var == "aquaculture NJ gear raised oysters sold") %>% 
  add_row(., Time = 2016,
          Value = 2029500,
          Var = "aquaculture NJ gear raised oysters sold",
          EPU = "MAB",
          Units = "N")

#MD data for 2017 is not available at this time



```

