---
title: "Environmental and LTL Indicators GB and GOM"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    theme: lumen
    self_contained: false
    lib_dir: libs
---

```{r setup, include=FALSE}
image.dir<- here::here("docs/images")
#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center') #allows for inserting R code into captions

#Plotting and data libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(ecodata)
library(here)
library(kableExtra)
library(ggrepel)
library(stringr)
library(patchwork)
library(grid)
library(heatwaveR)

#General inline text input for report

#Council
council <- "New England Fishery Management Council"
council_abbr <- "NEFMC"

#Region identifiers
epu <- "Geroges Bank"
epu_abbr <- "GB"
region <- "New England"
region_abbr <- "NE" #Some commercial data organized by "MA" or "NE" regions, not by EPU 

############################# GIS SETUP ######################################################

#GIS libraries
library(sf)
library(rgdal)
library(raster)
#CRS
crs <- "+proj=longlat +lat_1=35 +lat_2=45 +lat_0=40 +lon_0=-77 +x_0=0 +y_0=0 +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"

#EPU shapefile
epu_sf <- ecodata::epu_sf %>% 
  filter(EPU %in% c("MAB","GB","GOM"))

#Map line parameters
map.lwd <- 0.4

# Set lat/lon window for maps
xmin = -77
xmax = -65
ymin = 36
ymax = 45
xlims <- c(xmin, xmax)
ylims <- c(ymin, ymax)

#Time series constants
shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
feeding.guilds <- c("Apex Predator","Piscivore","Planktivore","Benthivore","Benthos")
x.shade.min <- 2012
x.shade.max <- 2022
#Function for custom ggplot facet labels
label <- function(variable,value){
  return(facet_names[value])
}

```


Trend lines are shown when slope is significantly different from 0 at the p < 0.05 level. An orange line signifies an overall positive trend, and purple signifies a negative trend. Note that in the final report we will only test for trend when N >= 30. However, I have relaxed that requirement for the purposes of this document so that trends are highlighted when N >= 20. **This means that some trends shown here will not be present in the final document**. Dashed lines represent mean values of time series unless the indicator is an anomaly, in which case the dashed line is equal to 0. Shaded regions indicate the past ten years. If there are no new data for 2018, the shaded region will still cover this time period.


## Gulf of Maine & Georges Bank

### Ocean temperature {.tabset .tabset-fade}

#### SST 

```{r seasonal-sst-anomaly-gridded, fig.width=8, fig.height = 8, fig.cap="GB and GOM 2022 seasonal sea surface temperature spatial anomalies. "}
p <- plot_sst()
p
```

```{r seasonal-oisst-anom, fig.width = 8, fig.cap="GB and GOM seasonal sea surface temperature anomaly time series."}
p <- plot_sst_anomaly(shadedRegion=c(2012,2022))
p
```



#### SST transition dates

```{r transdates,  fig.cap="Annual sea surface transition dates"}
p <- plot_sst_transition_dates()
p 

```

```{r sumlength,  fig.cap="Annual length of summer transition dates"}
p <- plot_sst_transition_dates_length()
p  
```

#### Bottom temperature

```{r seasonal-bt-anomaly-gridded, fig.width=8, fig.height = 8, fig.cap="GB and GOM 2022 seasonal bottom temperature spatial anomalies. "}
p <- plot_bottom_temp_anomaly_space()
p

```

```{r seasonal-bt-anom, fig.width = 8, fig.cap="GB and GOM seasonal Bottom Temperature anomaly time series."}
p <- plot_bottom_temp_anomaly_time()
p

```



```{r bottom-temp-insituglorys, fig.cap="GB and GOM annual bottom temperature anomalies."}

gb_bottomtemp <- plot_bottom_temp_anomaly_glorys(region="GB")

gom_bottomtemp <- plot_bottom_temp_anomaly_glorys(region="GOM")


plot_row<-cowplot::plot_grid(gom_bottomtemp,
                             gb_bottomtemp, ncol = 2)

title <- cowplot::ggdraw() +
  cowplot::draw_label(
    "Bottom Temperature Anomaly",
      x = 0,
      hjust = 0
    ) +
    ggplot2::theme(
      # add margin on the left of the drawing canvas,
      # so title is aligned with left edge of first plot
      plot.margin = margin(0, 0, 0,1)
    )


  p <- cowplot::plot_grid(title, plot_row, ncol = 1,
                     rel_heights = c(0.1, 1))



p

```


```{r bottom-temp-glorys, fig.width = 6, fig.asp = .55, fig.cap="Annual bottom temperature in the New England (black = Paula's, red = GLORYS)."}
# ecodata::bottom_temp_glorys %>% 
#   dplyr::filter(!EPU == "MAB", 
#                 !EPU == "SS")%>%
#   dplyr::group_by(EPU) %>% 
#   dplyr::mutate(hline = mean(Value)) %>% 
# ggplot2::ggplot() +
#   ggplot2::annotate("rect", fill = shade.fill, alpha = shade.alpha,
#       xmin = x.shade.min , xmax = x.shade.max,
#       ymin = -Inf, ymax = Inf) +
#   ggplot2::geom_line(aes(x = Time, y = Value)) +
#   #ecodata::geom_gls(aes(x = Time, y = Value)) +
#   ggplot2::geom_point(aes(x = Time, y = Value), size = 1) +
#   ggplot2::ylab("Temperature (C)") +
#   ggplot2::xlab(element_blank())+
#   
#   ggplot2::ggtitle("Bottom temperature GLORYS anomaly") +
#   ggplot2::scale_x_continuous(expand = c(0.01, 0.01)) +
#   ggplot2::theme(strip.text=element_text(hjust=0),
#         plot.title = element_text(size = 12))+
#   ggplot2::geom_hline(aes(yintercept = hline),
#            size = hline.size,
#            alpha = hline.alpha,
#            linetype = hline.lty) +
#   ggplot2::facet_wrap(~EPU)+
#   ecodata::theme_ts()

p <- plot_bottom_temp_anomaly_survey(region = "GB")
p
p <- plot_bottom_temp_anomaly_survey(region = "GOM")
p



```

#### GOM Slopewater Proportions

```{r slopewater, fig.width=5, fig.asp = 0.55, fig.cap="Proportion of warm slope water (WSW) and Labrador slope water (LSLW) entering the Gulf of Maine through the Northeast Channel."}

sw.df <- ecodata::slopewater %>% 
  dplyr::mutate(Var, Var = plyr::mapvalues(Var, from = c("WSW proportion ne channel",
                                                  "LSLW proportion ne channel"),
                                    to = c("WSW","LSW"))) %>% 
  dplyr::rename(Flavor  = Var) %>% 
  dplyr::group_by(Flavor) %>% 
  dplyr::mutate(hline = mean(Value, na.rm = TRUE)) 

sw.df$Origin <- factor(sw.df$Flavor, levels = c("WSW","LSW"))

ggplot2::ggplot(data = sw.df) +
  ggplot2::geom_line(aes(x = Time, y = Value, color = Origin))+
  ggplot2::geom_point(aes(x = Time, y = Value, color = Origin)) +
  #ecodata::geom_lm(aes(x = Time, y = Value, color = Origin))+
  ggplot2::ylab("Percent of Total Slopewater") +
  ggplot2::xlab(element_blank())+
  ggplot2::ggtitle("Slopewater Proportions in NE Channel")+
  ggplot2::scale_x_continuous(expand = c(0.01, 0.01))+
  ggplot2::geom_hline(aes(yintercept = hline, color = Origin),
           size = hline.size, alpha = hline.alpha,
           linetype = hline.lty)+
  ecodata::theme_ts() +
  ecodata::theme_title()

```


### Chlorophyll and Primary Productivity {.tabset .tabset-fade}

#### Primary production

```{r pp-monthly, fig.width=8, fig.height = 8, fig.cap="Monthly primary production trends show the annual cycle (i.e. the peak during the summer months) and the changes over time for each month."}
out_pp <- ecodata::chl_pp %>% 
  dplyr::filter(EPU %in% c("GOM","GB"),
         stringr::str_detect(Var, "MONTHLY_PPD_MEDIAN")) %>% 
  tidyr::separate(.,Time, into = c("Cat", "Time2"), sep = "_") %>% 
  tidyr::separate(.,Time2, into = c("Year", "Month"), sep = 4) %>% 
  dplyr::mutate(Month = plyr::mapvalues(Month, from = c("01","02","03","04","05","06",
                                                   "07","08","09","10","11","12"),
                                   to = c(month.abb))) %>% 
  dplyr::filter(!Value == "NA") %>% 
  dplyr::group_by(EPU, Month) %>% 
  dplyr::mutate(hline = mean(Value))
out_pp$Month <- factor(out_pp$Month, levels = month.abb)

pp_cci_gom <- out_pp %>% 
  dplyr::filter(EPU == "GOM") %>% 
 ggplot2::ggplot() +
   # geom_gls(aes(x = Year, y = Value, group = Month))+
    #ecodata::geom_lm(aes(x = Year, y = Value, group = Month))+
    ggplot2::geom_point(aes(x = Year, y = Value, group = Month)) +
    ggplot2::geom_line(aes(x = Year, y = Value, group = Month)) +
    ggplot2::scale_x_discrete(name = "", breaks = seq(min(out_pp$Year),max(out_pp$Year),10)) +  
    ggplot2::facet_wrap(Month~., ncol = 12) +
    ggplot2::ggtitle("GOM Monthly median PPD") +
    ggplot2::ylab(expression("PP (gC m"^-2*" d"^-1*")")) +
    ggplot2::geom_hline(aes(yintercept = hline,
                     group = Month),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
    ecodata::theme_facet() +
    ggplot2::theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm"))+
  ecodata::theme_title()
 
 pp_cci_gb <-out_pp %>% 
  dplyr::filter(EPU == "GB") %>% 
 ggplot2::ggplot() +
   # geom_gls(aes(x = Year, y = Value, group = Month))+
    #ecodata::geom_lm(aes(x = Year, y = Value, group = Month))+
    ggplot2::geom_point(aes(x = Year, y = Value, group = Month)) +
    ggplot2::geom_line(aes(x = Year, y = Value, group = Month)) +
    ggplot2::scale_x_discrete(name = "", breaks = seq(min(out_pp$Year),max(out_pp$Year),10)) +  
    ggplot2::facet_wrap(Month~., ncol = 12) +
    ggplot2::ggtitle("GB Monthly median PPD") +
    ggplot2::ylab(expression("PP (gC m"^-2*" d"^-1*")")) +
    ggplot2::geom_hline(aes(yintercept = hline,
                     group = Month),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
    ecodata::theme_facet() +
    ggplot2::theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm"))+
  ecodata::theme_title()
 
 pp_cci_gb + pp_cci_gom + patchwork::plot_layout(ncol = 1)
 
```


```{r chl-monthly, fig.width=8, fig.height = 8, fig.cap="Monthly chlorophyll trends show the annual cycle (i.e. the peak during the summer months) and the changes over time for each month."}
out_pp <- ecodata::chl_pp %>% 
  dplyr::filter(EPU %in% c("GOM","GB"),
         stringr::str_detect(Var, "MONTHLY_CHLOR_A_MEDIAN")) %>% 
   tidyr::separate(.,Time, into = c("Cat", "Time2"), sep = "_") %>% 
  tidyr::separate(.,Time2, into = c("Year", "Month"), sep = 4) %>% 
  dplyr::mutate(Month = plyr::mapvalues(Month, from = c("01","02","03","04","05","06",
                                                   "07","08","09","10","11","12"),
                                   to = c(month.abb))) %>% 
  dplyr::filter(!Value == "NA") %>% 
  dplyr::group_by(EPU, Month) %>% 
  dplyr::mutate(hline = mean(Value))
out_pp$Month <- factor(out_pp$Month, levels = month.abb)

pp_cci_gom <- out_pp %>% 
  dplyr::filter(EPU == "GOM") %>% 
 ggplot2::ggplot() +
    #ecodata::geom_lm(aes(x = Year, y = Value, group = Month))+
    ggplot2::geom_point(aes(x = Year, y = Value, group = Month)) +
    ggplot2::geom_line(aes(x = Year, y = Value, group = Month)) +
    ggplot2::scale_x_discrete(name = "", breaks = seq(min(out_pp$Year),max(out_pp$Year),10)) +  
    ggplot2::facet_wrap(Month~., ncol = 12) +
    ggplot2::ggtitle("GOM Monthly median CHLA") +
    ggplot2::ylab(expression("CHL (mg m"^-3*")")) +
    ggplot2::geom_hline(aes(yintercept = hline,
                     group = Month),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
    ecodata::theme_facet() +
    ggplot2::theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm"))+
  ecodata::theme_title()
 
 pp_cci_gb <-out_pp %>% 
  dplyr::filter(EPU == "GB") %>% 
 ggplot2::ggplot() +
    #ecodata::geom_lm(aes(x = Year, y = Value, group = Month))+
    ggplot2::geom_point(aes(x = Year, y = Value, group = Month)) +
    ggplot2::geom_line(aes(x = Year, y = Value, group = Month)) +
    ggplot2::scale_x_discrete(name = "", breaks = seq(min(out_pp$Year),max(out_pp$Year),10)) +  
    ggplot2::facet_wrap(Month~., ncol = 12) +
    ggplot2::ggtitle("GB Monthly median CHLA") +
    ggplot2::ylab(expression("CHL (mg m"^-3*")")) +
    ggplot2::geom_hline(aes(yintercept = hline,
                     group = Month),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
    ecodata::theme_facet() +
    ggplot2::theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm"))+
  ecodata::theme_title()
 
 pp_cci_gb + pp_cci_gom + patchwork::plot_layout(ncol = 1)
 
```

#### Seasonal Chlorophyll *a* & Primary Production

```{r pp-weekly, fig.width = 8, fig.cap = "Weekly chlorophyll concentrations and primary productivity for 2021 in Georges Bank and Gulf of Maine are shown by the colored lines in the above figures. The long-term mean is shown in black and shading indicates +/- 1 sample SD."}
interp_chl_pp <- function(epu, year = 2022, Variable){
  out <- ecodata::chl_pp %>%
    dplyr::filter(stringr::str_detect(Var,Variable),
           EPU == epu) %>%
    tidyr::separate(.,Time, into = c("Cat", "Time2"), sep = "_") %>%
    tidyr::separate(.,Time2, into = c("Year", "Week"), sep = 4)%>%
    dplyr::filter(Year == year) %>%
    dplyr::group_by(EPU) %>%
    dplyr::mutate(Time = 1:length(Year))

  ltm_out <- ecodata::chl_pp %>%
    dplyr::filter(stringr::str_detect(Var,Variable),
           EPU == epu) %>%
    tidyr::separate(.,Time, into = c("Cat", "Time2"), sep = "_") %>%
    tidyr::separate(.,Time2, into = c("Year", "Week"), sep = 4) %>%
    dplyr::group_by(Week) %>%
    dplyr::summarise(LTM = mean(Value, na.rm = T),
                     SD = sd(Value, na.rm = T)) %>%
    dplyr::mutate(Time = 1:length(Week),
           sd.low = LTM - SD,
           sd.high = LTM + SD) %>%
    dplyr::left_join(.,out, by = c("Time")) %>%
    dplyr::mutate(status = ifelse(Value < sd.high & Value > sd.low, "near_mean",
                           ifelse(Value > sd.high, "high",
                                  ifelse(Value < sd.low,"low",NA))),
           group = "PLOT")

  return(ltm_out)
}



GB_ppd_weekly <- interp_chl_pp(epu = "GB",  Variable = "WEEKLY_PPD_MEDIAN")
GOM_ppd_weekly <- interp_chl_pp(epu = "GOM", Variable = "WEEKLY_PPD_MEDIAN")
ne_ppd_weekly<-rbind(GB_ppd_weekly, GOM_ppd_weekly) %>% 
  dplyr::filter(!Value == "NA")

 #ne_pp_early<-ne_ppd_weekly %>% filter(Time<=26)
 #ne_pp_late<-ne_ppd_weekly %>% filter(Time >=26)

ne_ppd <- ggplot2::ggplot(data = ne_ppd_weekly) +
  ggplot2::geom_line(aes(x = Time, y = LTM)) +
  ggplot2::geom_ribbon(aes(x = Time, ymin = pmax(sd.low,0), ymax = sd.high),
              alpha = 0.1,
              fill = "grey1") +
  ggplot2::geom_line(aes(x = Time, y = Value),
            size = 1,color = "#33a02c") +
  # ggplot2::geom_line(data = ne_pp_late, aes(x = Time, y = LTM)) +
  # ggplot2::geom_ribbon(data = ne_pp_late,aes(x = Time, ymin = pmax(sd.low,0), ymax = sd.high),
  #             alpha = 0.1,
  #             fill = "grey1") +
  # ggplot2::geom_line(data = ne_pp_late,aes(x = Time, y = Value),
  #           size = 1,color = "#33a02c", linetype = "dashed") +
  ggplot2::ggtitle(expression("Primary Production")) +
  guides(color = F) +
  ggplot2::facet_wrap(EPU~., ncol = 2)+
  ggplot2::xlab("")+
  ggplot2::ylab("gC m^-2 d^-1") +
  ggplot2::scale_x_continuous(breaks = seq(1,52,10),
                   labels = c("Jan.","Mar.","May","July","Oct.","Dec."),
                   expand = c(0.01,0.01)) +
  ggplot2::scale_color_manual(values = c("#ef8a62","#2c7fb8","#a1d99b"))+
  ecodata::theme_ts()+
  ecodata::theme_title()+
  ecodata::theme_facet()+
  ggplot2::theme(panel.spacing = unit(1, "lines"))


ne_ppd


```

```{r chl-weekly, fig.width = 8, fig.cap = "Weekly chlorophyll concentrations and primary productivity for 2021 in Georges Bank and Gulf of Maine are shown by the colored lines in the above figures. The long-term mean is shown in black and shading indicates +/- 1 sample SD..."}
interp_chl_pp <- function(epu, year = 2022, Variable){
  out <- ecodata::chl_pp %>%
    dplyr::filter(stringr::str_detect(Var,Variable),
           EPU == epu) %>%
    tidyr::separate(.,Time, into = c("Cat", "Time2"), sep = "_") %>%
    tidyr::separate(.,Time2, into = c("Year", "Week"), sep = 4)%>%
    dplyr::filter(Year == year) %>%
    dplyr::group_by(EPU) %>%
    dplyr::mutate(Time = 1:length(Year))

  ltm_out <- ecodata::chl_pp %>%
    dplyr::filter(stringr::str_detect(Var,Variable),
           EPU == epu) %>%
    tidyr::separate(.,Time, into = c("Cat", "Time2"), sep = "_") %>%
    tidyr::separate(.,Time2, into = c("Year", "Week"), sep = 4) %>%
    dplyr::group_by(Week) %>%
    dplyr::summarise(LTM = mean(Value, na.rm = T),
                     SD = sd(Value, na.rm = T)) %>%
    dplyr::mutate(Time = 1:length(Week),
           sd.low = LTM - SD,
           sd.high = LTM + SD) %>%
    dplyr::left_join(.,out, by = c("Time")) %>%
    dplyr::mutate(status = ifelse(Value < sd.high & Value > sd.low, "near_mean",
                           ifelse(Value > sd.high, "high",
                                  ifelse(Value < sd.low,"low",NA))),
           group = "PLOT")

  return(ltm_out)
}

GB_chl_weekly <- interp_chl_pp(epu = "GB", Variable = "WEEKLY_CHLOR_A_MEDIAN")
GOM_chl_weekly <- interp_chl_pp(epu = "GOM", Variable = "WEEKLY_CHLOR_A_MEDIAN")
ne_chl_weekly<-rbind(GB_chl_weekly, GOM_chl_weekly)%>% 
  dplyr::filter(!Value == "NA")
# 
 # ne_early<-ne_chl_weekly %>% filter(Time <=26)
 # ne_late<-ne_chl_weekly %>% filter(Time >=26)

ne_chl <- ggplot2::ggplot(data = ne_chl_weekly) +
  ggplot2::geom_line(aes(x = Time, y = LTM)) +
  ggplot2::geom_ribbon(aes(x = Time, ymin = pmax(sd.low,0), ymax = sd.high),
              alpha = 0.1,
              fill = "grey1") +
  ggplot2::geom_line(aes(x = Time, y = Value),
            size = 1,color = "#33a02c") +
   # ggplot2::geom_line(data = ne_late, aes(x = Time, y = LTM)) +
   # ggplot2::geom_ribbon(data = ne_late,aes(x = Time, ymin = pmax(sd.low,0), ymax = sd.high),
   #             alpha = 0.1,
   #             fill = "grey1") +
   # ggplot2::geom_line(data = ne_late,aes(x = Time, y = Value),
   #           size = 1,color = "#33a02c", linetype = "dashed") +
  ggplot2::ggtitle(expression("Chlorophyll"~italic(a)~"")) +
  ggplot2::ylim(c(0, 3))+
  ggplot2::facet_wrap(EPU~., ncol = 2)+
  ggplot2::guides(color = F) +
  ggplot2::xlab("")+
  ggplot2::ylab("mg m^-3") +
  ggplot2::scale_x_continuous(breaks = seq(1,52,10),
                   labels = c("Jan.","Mar.","May","July","Oct.","Dec."),
                   expand = c(0.01,0.01)) +
  ggplot2::scale_color_manual(values = c("#ef8a62","#2c7fb8","#a1d99b"))+
  ecodata::theme_ts()+
  ecodata::theme_title()+
  ecodata::theme_facet()+
  ggplot2::theme(panel.spacing = unit(1, "lines"))

ne_chl

```


#### Phytoplankton Size class

```{r weekly-phyto-size, eval = T}

month <- seq(as.Date("2022-01-01"), 
             as.Date("2022-12-31"), 
             by = "1 month")
month_numeric <- lubridate::yday(month) / 365 * 52 + 1
month_label <- lubridate::month(month, label = TRUE)

phyto_year_nano<- ecodata::phyto_size %>% 
  dplyr::filter(EPU == c("GB"),
                Var %in% c("WEEKLY_PSC_FNANO_MEDIAN", 
                           "WEEKLY_PSC_FMICRO_MEDIAN")) %>% 
  mutate(Value = as.numeric(Value)) %>% 
  # dplyr::mutate(filter_me = 1:2600) %>% 
  # dplyr::filter(!filter_me == 2463,
  #               !filter_me == 1214) %>%
  # dplyr::select(!filter_me) %>% 
  tidyr::pivot_wider(names_from = "Var", values_from = "Value") %>%
  dplyr::mutate(nano = as.numeric(WEEKLY_PSC_FNANO_MEDIAN) +
                  as.numeric(WEEKLY_PSC_FMICRO_MEDIAN)) %>% 
  tidyr::separate(.,Time, into = c("Cat", "WEEK"), sep = "_") %>%
  dplyr::mutate(year = stringr::str_sub(WEEK, 1,4), 
                wk = stringr::str_sub(WEEK, 5,6)) %>% 
  tidyr::pivot_longer(cols = c("nano"), 
                      names_to = "Var", values_to = "Value") %>% 
  dplyr::filter(year == 2022, 
                !Value == "NA") %>% 
  dplyr::mutate(Value = Value*100)


phyto_year_micro<- ecodata::phyto_size %>% 
  dplyr::filter(EPU == c("GB"),
                Var == c("WEEKLY_PSC_FMICRO_MEDIAN")) %>% 
  mutate(Value = as.numeric(Value)) %>% 
  tidyr::separate(.,Time, into = c("Cat", "WEEK"), sep = "_") %>%
  dplyr::mutate(year = stringr::str_sub(WEEK, 1,4), 
                wk = stringr::str_sub(WEEK, 5,6)) %>% 
  dplyr::filter(year == 2022, 
                !Value == "NA") %>% 
  dplyr::mutate(Value = Value*100)

out_phyto<-  ecodata::phyto_size %>% 
  dplyr::filter(EPU == c("GB"),
                stringr::str_detect(Var, ("CLIMATOLOGICAL_WEEK"))) %>% #, 
                #!Var == "CLIMATOLOGICAL_WEEK_PSC_PICO_MEDIAN",
                #!Var == "CLIMATOLOGICAL_WEEK_PSC_NANO_MEDIAN",
                #!Var == "CLIMATOLOGICAL_WEEK_PSC_MICRO_MEDIAN") %>% 
  #tidyr::pivot_longer(cols = Var, names_to = "Var2", values_to = "Vaue2" )
  tidyr::separate(Time, into = c("Cat", "WEEK", "Year1", "Year2"), sep = "_") %>% 
  dplyr::filter(!Value == "NA", 
                !Var == "CLIMATOLOGICAL_WEEK_CHLOR_A_MEDIAN")  %>%
  dplyr::mutate(Value = Value*100) 
  
  
p<-  ggplot2::ggplot() +
  geom_area(aes(x=as.numeric(out_phyto$WEEK), y=out_phyto$Value, 
                fill = factor(out_phyto$Var, c("CLIMATOLOGICAL_WEEK_PSC_FPICO_MEDIAN",
                                               "CLIMATOLOGICAL_WEEK_PSC_FNANO_MEDIAN",
                                               "CLIMATOLOGICAL_WEEK_PSC_FMICRO_MEDIAN"))), alpha=0.6)+
  #ggplot2::geom_point(data = chlor, aes(x = as.numeric(WEEK), y = Value)) +
  ggplot2::geom_line( aes(x = as.numeric(phyto_year_nano$wk), 
                          y = phyto_year_nano$Value), color = "#FC8D62", size = 1.5)+
  ggplot2::geom_line( aes(x = as.numeric(phyto_year_micro$wk), 
                          y = phyto_year_micro$Value), color = "#66C2A5", size = 1.5)+
  #ggplot2::facet_wrap(EPU~., ncol = 2)+
  ggplot2::ggtitle("Georges Bank") +
  ggplot2::ylab("Percent") +
  ggplot2::xlab(element_blank())+
  ecodata::theme_facet() +
  ggplot2::theme(axis.text.x = element_text(angle=45, hjust = 1),
                 panel.spacing = unit(.5, "lines"),
                 plot.margin = unit(c(0.1, 0, 0, 0), "cm"), 
                 legend.position= "none")+
  scale_fill_manual(values=c("#8DA0CB", "#FC8D62","#66C2A5"), name = "",
                    labels = c("Picoplankton",
                               "Nanoplankton", "Microplankton"))+
  #scale_y_continuous( name = "Phytoplankton Size Fraction", sec.axis = sec_axis(~.*2, name="Chlorophyll a (mg m^-3)"))+
  scale_x_continuous(breaks = month_numeric, 
                     labels = month_label)+
  ecodata::theme_title()

month <- seq(as.Date("2021-01-01"),
             as.Date("2021-12-01"),
             by = "1 month")
month_numeric <- lubridate::yday(month) / 365 * 52 + 1
month_label <- lubridate::month(month, label = TRUE)

### GOM
phyto_year_nano2<- ecodata::phyto_size %>% 
  dplyr::filter(EPU == c("GOM"),
                Var %in% c("WEEKLY_PSC_FNANO_MEDIAN", 
                           "WEEKLY_PSC_FMICRO_MEDIAN")) %>% 
  mutate(Value = as.numeric(Value)) %>% 
  # dplyr::mutate(filter_me = 1:2600) %>% 
  # dplyr::filter(!filter_me == 2463,
  #               !filter_me == 1214) %>%
  # dplyr::select(!filter_me) %>% 
  tidyr::pivot_wider(names_from = "Var", values_from = "Value") %>%
  dplyr::mutate(nano = as.numeric(WEEKLY_PSC_FNANO_MEDIAN) +
                  as.numeric(WEEKLY_PSC_FMICRO_MEDIAN)) %>% 
  tidyr::separate(.,Time, into = c("Cat", "WEEK"), sep = "_") %>%
  dplyr::mutate(year = stringr::str_sub(WEEK, 1,4), 
                wk = stringr::str_sub(WEEK, 5,6)) %>% 
  tidyr::pivot_longer(cols = c("nano"), 
                      names_to = "Var", values_to = "Value") %>% 
  dplyr::filter(year == 2022, 
                !Value == "NA") %>% 
  dplyr::mutate(Value = Value*100)


phyto_year_micro2<- ecodata::phyto_size %>% 
  dplyr::filter(EPU == c("GOM"),
                Var == c("WEEKLY_PSC_FMICRO_MEDIAN")) %>% 
  mutate(Value = as.numeric(Value)) %>% 
  tidyr::separate(.,Time, into = c("Cat", "WEEK"), sep = "_") %>%
  dplyr::mutate(year = stringr::str_sub(WEEK, 1,4), 
                wk = stringr::str_sub(WEEK, 5,6)) %>% 
  dplyr::filter(year == 2022, 
                !Value == "NA") %>% 
  dplyr::mutate(Value = Value*100)

out_phyto2<-  ecodata::phyto_size %>% 
  dplyr::filter(EPU == c("GOM"),
                stringr::str_detect(Var, ("CLIMATOLOGICAL_WEEK"))) %>% #, 
                #!Var == "CLIMATOLOGICAL_WEEK_PSC_PICO_MEDIAN",
                #!Var == "CLIMATOLOGICAL_WEEK_PSC_NANO_MEDIAN",
                #!Var == "CLIMATOLOGICAL_WEEK_PSC_MICRO_MEDIAN") %>% 
  #tidyr::pivot_longer(cols = Var, names_to = "Var2", values_to = "Vaue2" )
  tidyr::separate(Time, into = c("Cat", "WEEK", "Year1", "Year2"), sep = "_") %>% 
  dplyr::filter(!Value == "NA", 
                !Var == "CLIMATOLOGICAL_WEEK_CHLOR_A_MEDIAN")  %>%
  dplyr::mutate(Value = Value*100) 
  
  
p2<-  ggplot2::ggplot() +
  geom_area(aes(x=as.numeric(out_phyto2$WEEK), y=out_phyto2$Value, 
                fill = factor(out_phyto2$Var, c("CLIMATOLOGICAL_WEEK_PSC_FPICO_MEDIAN",
                                               "CLIMATOLOGICAL_WEEK_PSC_FNANO_MEDIAN",
                                               "CLIMATOLOGICAL_WEEK_PSC_FMICRO_MEDIAN"))), alpha=0.6)+
  #ggplot2::geom_point(data = chlor, aes(x = as.numeric(WEEK), y = Value)) +
  ggplot2::geom_line( aes(x = as.numeric(phyto_year_nano2$wk), 
                          y = phyto_year_nano2$Value), color = "#FC8D62", size = 1.5)+
  ggplot2::geom_line( aes(x = as.numeric(phyto_year_micro2$wk), 
                          y = phyto_year_micro2$Value), color = "#66C2A5", size = 1.5)+
  #ggplot2::facet_wrap(EPU~., ncol = 2)+
  ggplot2::ggtitle("Gulf of Maine") +
  ggplot2::ylab(element_blank()) +
  ggplot2::xlab(element_blank())+
  ecodata::theme_facet() +
  ggplot2::theme(axis.text.x = element_text(angle=45, hjust = 1),
                 panel.spacing = unit(.5, "lines"),
                 plot.margin = unit(c(0.1, 0, 0, 0), "cm"), 
                 legend.position= "none", 
  axis.text.y = element_blank(), 
  axis.ticks.y = element_blank())+
  scale_fill_manual(values=c("#8DA0CB", "#FC8D62","#66C2A5"), name = "",
                    labels = c("Picoplankton",
                               "Nanoplankton", "Microplankton"))+
  #scale_y_continuous( name = "Phytoplankton Size Fraction", sec.axis = sec_axis(~.*2, name="Chlorophyll a (mg m^-3)"))+
  scale_x_continuous(breaks = month_numeric, 
                     labels = month_label)+
  ecodata::theme_title()


plot_row<- cowplot::plot_grid(p, p2, ncol = 2)
title <- cowplot::ggdraw() + 
  cowplot::draw_label(
    "Phytoplankton Size Class",
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0,1)
  )


plot_w_title <- cowplot::plot_grid(title, plot_row, ncol = 1, 
                   rel_heights = c(0.1, 1))
legend_b <- cowplot::get_legend(
  p + 
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")
)

cowplot::plot_grid(plot_w_title, legend_b, ncol = 1,rel_heights = c(1, 0.1) )
```



<!-- ```{r seasonal-phyto-size} -->

<!-- micro_chlor <- ecodata::phyto_size %>%  -->
<!--   dplyr::filter(EPU %in% c("GB", "GOM"), -->
<!--                 Var %in% c("CLIMATOLOGICAL_WEEK_MICRO_PERCENTAGE_MEDIAN",  -->
<!--                            "CLIMATOLOGICAL_WEEK_CHLOR_A_MEDIAN")) %>%  -->
<!--   tidyr::separate(.,Time, into = c("Cat", "WEEK", "Year1", "Year2"), sep = "_") %>%  -->
<!--   dplyr::filter(!Value == "NA") %>%  -->
<!--   dplyr::select(-Year1, -Year2, -Cat,  -Units) %>%  -->
<!--   tidyr::pivot_wider(id_cols= c(EPU, WEEK), names_from = Var, values_from = Value) %>%  -->
<!--   dplyr::mutate(Month = rep(1:4, each = 26)) %>%  -->
<!--   ggplot2::ggplot() + -->
<!--   ggplot2::geom_point( aes(x = CLIMATOLOGICAL_WEEK_MICRO_PERCENTAGE_MEDIAN,  -->
<!--                            y = CLIMATOLOGICAL_WEEK_CHLOR_A_MEDIAN, color = factor(Month))) + -->


<!--   ggplot2::scale_color_discrete(name = "Season", labels=c("Jan-Mar", "Apr-Jun", "Jul-Sep", "Oct-Dec"))+ -->
<!--   ggplot2::ggtitle("") + -->
<!--   ggplot2::facet_wrap(~EPU)+ -->
<!--   ggplot2::ylab("Chlorophyll a (mg m^-3)") + -->
<!--   ggplot2::xlab("Microplankton % of Total Phytoplankton Composition ") + -->
<!--   ecodata::theme_facet()+ -->
<!--   ecodata::theme_title() -->
<!-- #micro_chlor -->

<!-- ``` -->


<!-- ```{r chl-monthly, fig.width = 8, fig.height=8, eval = F} -->
<!-- out_chl <- ecodata::chl_pp %>%  -->
<!--   dplyr::filter(EPU %in% c("GOM","GB"), -->
<!--          stringr::str_detect(Var, "MONTHLY_CHLOR_A_MEDIAN")) %>%  -->
<!--   tidyr::separate(.,Time, into = c("Year","Month"), sep = 4) %>%  -->
<!--   dplyr::mutate(Month = plyr::mapvalues(Month, from = c("01","02","03","04","05","06", -->
<!--                                                    "07","08","09","10","11","12"), -->
<!--                                    to = c(month.abb))) -->
<!-- out_chl$Month <- factor(out_chl$Month, levels = month.abb) -->


<!-- chl_cci_gom <-  -->
<!--   out_chl %>%  -->
<!--   dplyr::filter(EPU == "GOM") %>%  -->
<!--   ggplot2::ggplot() + -->
<!--     #geom_gls(aes(x = Year, y = Value, group = Month))+ -->
<!--     ggplot2::geom_point(aes(x = Year, y = Value, group = Month)) + -->
<!--     ggplot2::geom_line(aes(x = Year, y = Value, group = Month)) + -->
<!--     ggplot2::scale_x_discrete(name = "", breaks = seq(min(out_chl$Year),max(out_chl$Year),10)) +   -->
<!--     ggplot2::facet_wrap(Month~., ncol = 12) + -->
<!--     ggplot2::ggtitle("GOM Monthly median CHL (OC-CCI)") + -->
<!--     ggplot2::ylab(expression("CHL (mg m"^-3*")")) + -->
<!--     ecodata::theme_facet() + -->
<!--     ggplot2::theme(axis.text.x = element_text(angle=45, hjust = 1), -->
<!--           panel.spacing = unit(0.5, "lines"), -->
<!--           plot.margin = unit(c(0.1, 0, 0, 0), "cm")) -->

<!-- chl_cci_gb <-  -->
<!--   out_chl %>%  -->
<!--   dplyr::filter(EPU == "GB") %>%  -->
<!--   ggplot2::ggplot() + -->
<!--     #geom_gls(aes(x = Year, y = Value, group = Month))+ -->
<!--     ggplot2::geom_point(aes(x = Year, y = Value, group = Month)) + -->
<!--     ggplot2::geom_line(aes(x = Year, y = Value, group = Month)) + -->
<!--     ggplot2::scale_x_discrete(name = "", breaks = seq(min(out_chl$Year),max(out_chl$Year),10)) +   -->
<!--     ggplot2::facet_wrap(Month~., ncol = 12) + -->
<!--     ggplot2::ggtitle("GB Monthly median CHL (OC-CCI)") + -->
<!--     ggplot2::ylab(expression("CHL (mg m"^-3*")")) + -->
<!--     ecodata::theme_facet() + -->
<!--     ggplot2::theme(axis.text.x = element_text(angle=45, hjust = 1), -->
<!--           panel.spacing = unit(0.5, "lines"), -->
<!--           plot.margin = unit(c(0.1, 0, 0, 0), "cm")) -->
<!-- chl_cci_gb + chl_cci_gom + patchwork::plot_layout(ncol = 1) -->

<!-- ``` -->

