---
title: 'State of the Ecosystem: Environmental Indicator Processing & Visualization'
author: Ecosystems Dynamics and Assessment Branch
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    theme: lumen
---

## Introduction

The purpose of this report is to document State of the Ecosystem (SOE) **lower trophic level** indicator data processing. All R code used to process and visualize the following data sets is self-contained in the [Rmarkdown document](https://github.com/NOAA-EDAB/soe/blob/master/inst/Rmd/process_raw.Rmd) associated with this HTML file. To run and update data sets in this document, set the `save_clean` parameter in the set-up chunk to `TRUE`. Raw data for these indicators are available in the file directory `soe/inst/extdata`.

```{css, echo = FALSE}
pre code, pre, code {
  white-space: pre !important;
  overflow-y: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
  max-height: 400px;
}
```

```{r, message=FALSE, echo = F}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align='center')

#Required libraries
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(here)
library(zoo)
library(kableExtra)
library(sf)
library(rgdal)
library(raster)
library(sp)
library(gridExtra)
library(htmlwidgets)
library(corrplot)
library(tseries)
library(heatmaply)
library(flexdashboard)
#library(plotly)
library(ggiraph)


#Data directories
raw.dir <- here("inst","extdata") #raw data directory
clean.dir <- here("data") #output directory for cleaned data
gis.dir <- here("inst","extdata","gis")
sample.dir <- here("inst","extdata","sample")

#CRS
crs <- "+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"

#Write output to file
save_clean <- F

#Execute spatial processing (must be TRUE to write clean data to file). If FALSE, will load sample data from file for plotting example
spatial_processing <- F

#Create interactive time series
gg_int <- TRUE

#Function for custom ggplot facet labels
label <- function(variable,value){
  return(facet_names[value])
}

```

### EcoMon zooplankton data (OI) {.tabset .tabset-fade}

These data show estimated zooplankton abundances on the NE Shelf derived from Ecosystem Monitoring Program (EcoMon) sampling. EcoMon conducts shelf-wide bimonthly surveys of the Northeast Large Marine Ecosystem, collecting zooplankton and ichthyoplankton to a depth of 200 m using paired Bongo samplers with 333 $\mu$m mesh netting. Zooplankton abundance data were interpolated across sampling locations using ordinary kriging to create a complete field. Here we present abundance time series for three species: *Centropages typicus*, *Temora longicornis*, and *Pseudocalanus* spp. These data are processed in a similar manner to optimally interpolated ocean temperature and salinity data sets. More information about the source data and processing methods used to derive these data sets can be found [here](https://noaa-edab.github.io/ECSA/#sec:methodszoo). 


#### Processing

*Function definitions & spatial processing*

```{r zooplankton_oi_processing}

#Read in EPU shapefile (will be downsampled to match raster resolution)
epu <- readOGR(file.path(gis.dir, "EPU_Extended.shp"), verbose = F) 

#Write function to build EPU rasters
build_raster <- function(EPU){
  
  #Filter raster by EPU
  epu <- epu[epu$EPU == EPU,]
  
  #Build empty raster
  r1 <- raster::raster()
  e <- raster::extent(-75.950, -65.450, 35.650, 44.650)
  raster::extent(r1) <- e
  
  #fill with EPU polygon
  r1 <- raster::rasterize(epu, r1, field = 1, fun = mean)
  raster::crs(r1) <- NA
  
  #create raster to resample with
  r2 <- raster::raster(nrow = 90, ncol = 105)
  raster::extent(r2) <- e
  raster::crs(r2) <- NA
  
  #Downsample high res EPU raster to match data
  r.new <- raster::resample(r1, r2, method="bilinear")
  r.new[is.finite(r.new)] <- 1 
  
  return(r.new)
  
}

#Raster data
mab_rast <- build_raster("MAB")
gb_rast <- build_raster("GB")
gom_rast <- build_raster("GOM")

#Process ECSA data by EPU and create time series

epu_env <- function(variable, type = NULL, season, genus = NULL, epu){
  

  if(!is.null(type) & !variable %in% c("salinity","temperature")){
    stop('type only applicable for variables "salinity" and "temperature"
         as type = "bottom" or type = "surface"')
  } 
  
  if(!is.null(genus) & variable != "zooplankton"){
    stop('genus only applicable for variable "zooplankton"')
  } 
  
  #get compiled down-sampled raster of chosen strata from shapefile. 
  epumask.raster <- get(paste0(tolower(epu),"_rast"))
  
  #get bottom temp data and find mean for stock area--------------------------------------
  
  indir <- here("inst","extdata","gridded")
  
  if (variable == "salinity"){
    load(file.path(indir, paste0("sal_",type,"_",season,"_spdf.rdata")))
  } else if (variable == "temperature"){
    load(file.path(indir, paste0("temp_",type,"_",season,"_spdf.rdata")))
  } else if (variable == "chlorophyll"){
    load(file.path(indir, paste0("chl_",season,"_1997-2018.rdata")))
  } else if (variable == "zooplankton"){
    load(file.path(indir, paste0(genus,"_",season,"_zoo_1977-2016.rdata")))
  }
  

    #create null df to fill with results
    data = data.frame(array(NA,dim= c(raster::nlayers(ecsa_dat),5)))
    
    #loops through layers in raster brick
    for(i in 1:raster::nlayers(ecsa_dat)){
      #load raster by year
      
      #get file information from title
      layer_id <- stringr::str_extract(names(ecsa_dat)[[i]], "\\d.*")
      layer_id <- stringr::str_split(layer_id, "_")
      data[i,1] <- layer_id[[1]][[1]]
      data[i,2] <- layer_id[[1]][[2]]
      data[i,3] <- layer_id[[1]][[3]]
      
      #trim to stock area
      masked.raster = ecsa_dat[[i]]*epumask.raster
      
      #find mean BT of stock area
      data[i,4] = raster::cellStats(masked.raster, stat='mean', na.rm=TRUE)
      data[i,5] = raster::cellStats(masked.raster, stat = 'sd', na.rm=TRUE)
      # 
      # if (layer_id[[1]][[1]] == "1995"){
      #   break
      # }
    }
    x <- as.numeric(data$X1)
    y.out <- data$X4
    y.sd <- data$X5
    
    sd.low <- y.out - y.sd
    sd.high <- y.out + y.sd
    
  # remove 
  if (variable == "zooplankton"){
    if (season == "spring"){
      y.out[x %in% c(1989, 1990, 1991, 1994)] <- NA
      sd.low[x %in% c(1989, 1990, 1991, 1994)] <- NA
      sd.high[x %in% c(1989, 1990, 1991, 1994)] <- NA
    } else if (season == "fall") {
      y.out[x %in% c(1989, 1990, 1992)] <- NA
      sd.low[x %in% c(1989, 1990, 1992)] <- NA
      sd.high[x %in% c(1989, 1990, 1992)] <- NA
    }
  }
  
  if(variable == "chlorophyll"){
    type <- ""
  } else if (variable == "zooplankton"){
    type <- genus
    variable <- "zoo"
  } 
  
  out <- data.frame(Var = paste(paste(type,variable),season),
                    Time = as.numeric(x),
                    Value = y.out,
                    sd.low = sd.low,
                    sd.high = sd.high,
                    Season = season,
                    epu = epu)
  
  out <- out[out$Time > 1968,]
  return(out)
  
}

```

*Data pull & aggregation*

```{r zoo_data_pull}
#MAB---------------------------------------------------------------------------------------------------
ctyp_fall_mab <- epu_env(variable = "zooplankton", genus = "centropages", season = "fall", epu = "MAB")
ctyp_spring_mab <- epu_env(variable = "zooplankton", genus = "centropages", season = "spring", epu = "MAB")

tlong_fall_mab <- epu_env(variable = "zooplankton", genus = "temora", season = "fall", epu = "MAB")
tlong_spring_mab <- epu_env(variable = "zooplankton", genus = "temora", season = "spring", epu = "MAB")

pseudo_fall_mab <- epu_env(variable = "zooplankton", genus = "pseudocalanus", season = "fall", epu = "MAB")
pseudo_spring_mab <- epu_env(variable = "zooplankton", genus = "pseudocalanus", season = "spring", epu = "MAB")


#GB---------------------------------------------------------------------------------------------------
ctyp_fall_gb <- epu_env(variable = "zooplankton", genus = "centropages", season = "fall", epu = "GB")
ctyp_spring_gb <- epu_env(variable = "zooplankton", genus = "centropages", season = "spring", epu = "GB")

tlong_fall_gb <- epu_env(variable = "zooplankton", genus = "temora", season = "fall", epu = "GB")
tlong_spring_gb <- epu_env(variable = "zooplankton", genus = "temora", season = "spring", epu = "GB")

pseudo_fall_gb <- epu_env(variable = "zooplankton", genus = "pseudocalanus", season = "fall", epu = "GB")
pseudo_spring_gb <- epu_env(variable = "zooplankton", genus = "pseudocalanus", season = "spring", epu = "GB")


#GOM---------------------------------------------------------------------------------------------------
ctyp_fall_gom <- epu_env(variable = "zooplankton", genus = "centropages", season = "fall", epu = "GOM")
ctyp_spring_gom <- epu_env(variable = "zooplankton", genus = "centropages", season = "spring", epu = "GOM")

tlong_fall_gom <- epu_env(variable = "zooplankton", genus = "temora", season = "fall", epu = "GOM")
tlong_spring_gom <- epu_env(variable = "zooplankton", genus = "temora", season = "spring", epu = "GOM")

pseudo_fall_gom <- epu_env(variable = "zooplankton", genus = "pseudocalanus", season = "fall", epu = "GOM")
pseudo_spring_gom <- epu_env(variable = "zooplankton", genus = "pseudocalanus", season = "spring", epu = "GOM")


```


#### Mid-Atlantic Bight

```{r mab_zoo_vis, echo = FALSE, warning = FALSE, message = FALSE}

zoo <- rbind(ctyp_fall_mab, ctyp_spring_mab, tlong_fall_mab, tlong_spring_mab, pseudo_fall_mab, pseudo_spring_mab)

xmin <- min(zoo$Time)

facet_names <- list(
  'centropages zoo spring'=expression(paste(italic("Centropages"), "spring")),
  'centropages zoo fall'=expression(paste(italic("Centropages"), "fall")),
  'temora zoo spring'=expression(paste(italic("Temora"), "spring")),
  'temora zoo fall' = expression(paste(italic("Temora"), "fall")),
  'temora zoo spring'=expression(paste(italic("Pseudocalanus"), "spring")),
  'temora zoo fall' = expression(paste(italic("Pseudocalanus"), "fall")))


zoo_plt <- ggplot(data = zoo, aes(x = Time, y = Value)) +
    ylab(expression("Abundance log num m"^-3*"")) +
    xlab("Year") +
    xlim(xmin, NA) +
    geom_line() +
    {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time))
    else geom_point()}+
    facet_wrap(Var ~ ., nrow = 3, ncol = 2, labeller = label) +
    theme_bw() +
    theme(strip.background = element_blank()) +
    annotate("text", label = c("","","",
                               "","",""), x = xmin, y = Inf, vjust = 1.5, size = 5)

if (gg_int){
  ggiraph(code = print(zoo_plt), width = 0.9, 
         hover_css = "fill:red;r:3pt;")
  } else {
  zoo_plt
  }


```

#### Georges Bank

```{r gb_zoo_vis, echo = FALSE, warning = FALSE, message = FALSE}

zoo <- rbind(ctyp_fall_gb, ctyp_spring_gb, tlong_fall_gb, tlong_spring_gb, pseudo_fall_gb, pseudo_spring_gb)

xmin <- min(zoo$Time)

zoo_plt <- ggplot(data = zoo, aes(x = Time, y = Value)) +
    ylab(expression("Abundance log num m"^-3*"")) +
    xlab("Year") +
    xlim(xmin, NA) +
    geom_line() +
    {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time))
    else geom_point()}+
    facet_wrap(Var ~ ., nrow = 3, ncol = 2, labeller = label) +
    theme_bw() +
    theme(strip.background = element_blank()) +
    annotate("text", label = c("","","",
                               "","",""), x = xmin, y = Inf, vjust = 1.5, size = 5)

if (gg_int){
  ggiraph(code = print(zoo_plt), width = 0.9, 
         hover_css = "fill:red;r:3pt;")
  } else {
  zoo_plt
  }

```


#### Gulf of Maine

```{r gom_zoo_vis, echo = FALSE, warning=FALSE, message = FALSE}

zoo <- rbind(ctyp_fall_gom, ctyp_spring_gom, tlong_fall_gom, tlong_spring_gom, pseudo_fall_gom, pseudo_spring_gom)

xmin <- min(zoo$Time)

zoo_plt <- ggplot(data = zoo, aes(x = Time, y = Value)) +
    ylab(expression("Abundance log num m"^-3*"")) +
    xlab("Year") +
    xlim(xmin, NA) +
    geom_line() +
    {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time))
    else geom_point()}+
    facet_wrap(Var ~ ., nrow = 3, ncol = 2, labeller = label) +
    theme_bw() +
    theme(strip.background = element_blank()) +
    annotate("text", label = c("","","",
                               "","",""), x = xmin, y = Inf, vjust = 1.5, size = 5)

if (gg_int){
  ggiraph(code = print(zoo_plt), width = 0.9, 
         hover_css = "fill:red;r:3pt;")
  } else {
  zoo_plt
  }

```


### EcoMon nutrient data {.tabset .tabset-fade}

These data include nutrient concentrations, temperature, salinity, density, and dissolved oxygen data sampled via CTD profiles on Ecosystem Monitoring (EcoMon) cruises between 11/3/2009 - 10/19/2016. More metadata are available [here](https://www.nodc.noaa.gov/oads/data/0127524.xml).  
</br>
<div align="center">
```{r EcoMon_vars, echo = FALSE}

fname <- "EcoMon Nutrient Data Through June 2018.csv"

vars <- data.frame(Variable = c("Cruise identifier","Cruise identifier","Station number",
                                "CTD cast number","Sample bottle number","Sample date",
                                "Sample time","Latitude","Longitude","Depth of station",
                                "Depth of sample","Water pressure","Water temperature",
                                "Water salinity","Potential density at surface pressure",
                                "Dissolved oxygen","Silicic acid concentration",
                                "Total nitrate and nitrite concentration","Ammonia concentration",
                                "Phosphate concentration","Dissolved oxygen"),
                   Names = c("EXPOCODE","Cruise_ID","STNNBR","CASTNO",
                             "BTLNBR","Date_UTC","Time_UTC",
                             "Latitude","Longitude","Depth_station",
                             "Depth_sampling","CTDPRS","CTDTEMP",
                             "CTDSAL","Sigma.Theta","CTDOXY",
                             "SILCAT","NITRIT+NITRAT","AMMMONIA",
                             "PHSPHT","CTDOXYMOL"),
                   Units = c("","","",
                             "","","MM/DD/YYYY",
                             "hh:mm","decimal degrees","decimal degrees",
                             "m","m","decibars","°C",
                             "PSS-78","kg m^-3^","mg L^-1^",
                             "$\\mu$M","$\\mu$M","$\\mu$M",
                             "$\\mu$M","$\\mu$mol kg^-1^"))

kable(vars,caption = paste0('Variables in "',fname,'"'))%>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center") %>%
  column_spec(2, width = "5cm") %>% 
  scroll_box(height = "400px")
```
</div>  
</br>
**Filename**: EcoMon Nutrient Data Through June 2018.csv  
**Contributor**: Chris Melrose (chris.melrose@noaa.gov)  

#### Processing

```{r EcoMon nutrients, message=FALSE, warning=FALSE}
d <- read.csv(file.path(raw.dir,"EcoMon Nutrient Data Through June 2018.csv"), stringsAsFactors = FALSE)

#Create data frame for mapping units to variable names
mapping <- data.frame(Units = as.character(d[1,]),
                      Var = as.character(names(d)))
mapping[mapping$Units == "" | mapping$Units == "NA",]$Units <- NA

#remove row with units
d <- slice(d,-1)

d1 <- d %>% 
  mutate(Time = Date_UTC) %>% #create Time variable
  dplyr::select(-Date_UTC,-Time_UTC) %>% #remove time, date
  gather(., Var, Value, -Latitude, -Longitude, -Time, -Depth_sampling, -Depth_station) %>% #turn wide to long while retaining lat/lon
  filter(!is.na(Value)) %>% #remove NA
  left_join(., mapping, by = c("Var")) %>% #join units 
  mutate(Longitude = as.numeric(Longitude),
         Latitude = as.numeric(Latitude),
         Time = mdy(Time)) %>% 
  filter(Latitude > 32, Latitude<50)


#Sanity check
# t1 <- d1[d1$Var == "CTDOXYMOL" ,]$Value
# t <-  d %>% slice(.,-1)
# t <- as.character(t$CTDOXYMOL)
# all(t == t1)

#Read in EPU shapefile
epu <- readOGR(file.path(gis.dir, "EPU_Extended.shp"), verbose = F) 
epu <- as(epu, "sf") #convert to sf object

if(spatial_processing){

  #Test maps
  #All EPUs
  #ggplot() + geom_sf(data = epu)
  
  #Scotian shelf
  # ss <- epu %>% filter(EPU == "SS")
  # ggplot() + geom_sf(data = ss)
  
  #get latitude and longitude for creating SpatialPointsDataFrame
  lat <-  as.numeric(d$Latitude)
  lon <- as.numeric(d$Longitude)
  coords <- data.frame(lon = lon,lat = lat)
  
  #create spdf
  spdf <- SpatialPointsDataFrame(coords = coords, data = coords,
                                 proj4string = CRS(crs))
  #convert to sf
  coords_sf <- st_as_sf(spdf) 
  
  #get intersection for mapping EPUs back to nutrient data
  epu_intersect <- st_intersection(epu, coords_sf)
  #plot(epu_intersect[epu_intersect$EPU == "MAB",])
  
  #Map back to nutrient data frame
  epu_df <- data.frame(Longitude = epu_intersect$lon,
                       Latitude = epu_intersect$lat,
                       EPU = epu_intersect$EPU)
  #join
  NE_LME_nutrients <- d1 %>% 
    left_join(.,epu_df, by = c("Latitude","Longitude"))
  
  #Select data for plotting 
  Nitr <- NE_LME_nutrients %>% filter(Var == "NITRIT.NITRAT")
  
  #Back to SOE format and specify bottom, mid-water, or surface sampling
  NE_LME_nutrients <- NE_LME_nutrients %>%
    dplyr::select(-Latitude, -Longitude) %>% 
    mutate(Value = as.numeric(Value),
           Depth_station = as.numeric(Depth_station),
           Depth_sampling = as.numeric(Depth_sampling)) %>% 
    mutate(bot_dif = Depth_station-Depth_sampling) %>% 
    mutate(surf_bot = ifelse(bot_dif <= 10, "bottom",
                           ifelse(bot_dif > 10 & Depth_sampling <= 5, "surface", "mid-water"))) %>% 
    filter(Value > 0, !is.na(EPU), !Var %in% c("BTLNBR","CASTNO","Depth_sampling",
                                             "Depth_station","STNNBR")) %>% 
    mutate(Var = paste(Var, surf_bot)) %>% 
    dplyr::select(Time, Var, Value, Units, EPU) %>% 
    group_by(EPU, Time = year(Time), Var, Units) %>% 
    dplyr::summarise(Value = mean(Value, na.rm = TRUE)) %>% 
    as.data.frame()
  
  if (save_clean){
    save(NE_LME_nutrients,file = file.path(clean.dir, "EcoMon_nutrients.Rdata"))
  }
} else {
  load(file.path(sample.dir,"sample_nutrients.Rdata"))
  load(file.path(clean.dir,"EcoMon_nutrients.Rdata"))
}

```

#### QA

```{r ecomon_plotting1, echo = F}
#Confirm transformation
ggplot() + 
  geom_sf(data = epu) +
  geom_point(data = Nitr, aes(x = Longitude, y = Latitude, color = EPU)) +
  ggtitle("Mapping EcoMon Nutrient Data to EPU") +
  theme_bw() 
```


#### Surface Nutrients

```{r surface_nutrients, echo = F}

N <- NE_LME_nutrients %>% 
  filter(Var == "NITRIT.NITRAT surface") %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Value, color = EPU)) +
  {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time,x = Time, y = Value, color = EPU))
    else geom_point()}+
  ylab(expression("Nitrit+Nitrat ("*mu*"mol kg"^-1*")"))+
  ggtitle("Nitrit+Nitrat")+
  theme_bw()+
  theme(strip.background = element_blank())+
  guides(colour=FALSE)

P <- NE_LME_nutrients %>% 
  filter(Var == "PHSPHT surface") %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Value, color = EPU)) +
  {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time,x = Time, y = Value, color = EPU))
    else geom_point()}+
  ylab(expression("Phosphate ("*mu*"mol kg"^-1*")"))+
  ggtitle("Phosphate")+
  theme_bw()+
  theme(strip.background = element_blank())+
  guides(colour=FALSE)

ammon <- NE_LME_nutrients %>% 
  filter(Var == "AMMONIA surface") %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Value, color = EPU)) +
  {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time,x = Time, y = Value, color = EPU))
    else geom_point()}+
  ylab(expression("Ammonia ("*mu*"mol kg"^-1*")"))+
  ggtitle("Ammonia")+
  theme_bw()+
  theme(strip.background = element_blank(),legend.position = c(0.85, 0.6),
        legend.key = element_rect(color="transparent"),
        legend.background = element_rect(fill="transparent"))

silcat <- NE_LME_nutrients %>% 
  filter(Var == "SILCAT surface") %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Value, color = EPU)) +
  {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time,x = Time, y = Value, color = EPU))
    else geom_point()}+
  ylab(expression("Silicates ("*mu*"mol kg"^-1*")"))+
  ggtitle("Silicates")+
  theme_bw()+
  guides(color= FALSE)

cow_nut <- cowplot::plot_grid(N, P,ammon ,silcat)

if (gg_int){
  ggiraph(code = print(cow_nut), width = 0.9, 
         hover_css = "fill:red;r:3pt;")
  } else {
    cow_nut
  }

```

#### Bottom Nutrients

```{r bottom_nutrients, echo = F}

N <- NE_LME_nutrients %>% 
  filter(Var == "NITRIT.NITRAT bottom") %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Value, color = EPU)) +
  {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time,x = Time, y = Value, color = EPU))
    else geom_point()}+
  ylab(expression("Nitrit+Nitrat ("*mu*"mol kg"^-1*")"))+
  ggtitle("Nitrit+Nitrat")+
  theme_bw()+
  theme(strip.background = element_blank())+
  guides(colour=FALSE)

P <- NE_LME_nutrients %>% 
  filter(Var == "PHSPHT bottom") %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Value, color = EPU)) +
  {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time,x = Time, y = Value, color = EPU))
    else geom_point()}+
  ylab(expression("Phosphate ("*mu*"mol kg"^-1*")"))+
  ggtitle("Phosphate")+
  theme_bw()+
  theme(strip.background = element_blank())+
  guides(colour=FALSE)

ammon <- NE_LME_nutrients %>% 
  filter(Var == "AMMONIA bottom") %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Value, color = EPU)) +
  {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time,x = Time, y = Value, color = EPU))
    else geom_point()}+
  ylab(expression("Ammonia ("*mu*"mol kg"^-1*")"))+
  ggtitle("Ammonia")+
  theme_bw()+
  theme(strip.background = element_blank(),legend.position = c(0.85, 0.6),
        legend.key = element_rect(color="transparent"),
        legend.background = element_rect(fill="transparent"))

silcat <- NE_LME_nutrients %>% 
  filter(Var == "SILCAT bottom") %>% 
  ggplot() +
  geom_line(aes(x = Time, y = Value, color = EPU)) +
  {if(gg_int) geom_point_interactive(aes(tooltip = round(Value,3), data_id = Time,x = Time, y = Value, color = EPU))
    else geom_point()}+
  ylab(expression("Silicates ("*mu*"mol kg"^-1*")"))+
  ggtitle("Silicates")+
  theme_bw()+
  guides(color= FALSE)

cow_nut <- cowplot::plot_grid(N, P,ammon ,silcat)

if (gg_int){
  ggiraph(code = print(cow_nut), width = 0.9, 
         hover_css = "fill:red;r:3pt;")
  } else {
    cow_nut
  }
```

