---
title: "Habitat Vulnerability Tables"
author: "Sean Lucey and Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(here); library(kableExtra); library(dplyr); library(readxl); library(data.table); library(DT);
library(stringr)
knitr::opts_chunk$set(echo = FALSE)

options(knitr.kable.NA = '')

```

```{css, echo=FALSE}

/*alter risk assessment table formatting*/
.risktable {
    border-collapse: collapse;
}
.risktable th {
    padding: 2px;
}
.risktable td {
    padding: 1px;
}
.risktable tbody tr:nth-of-type(even) {
  background-color:transparent;
}

```

#### Mid-Atlantic Habitat Vulnerability Table

This table aligns species by life stage (rows) with habitats (columns) with the climate vulnerability rankings for both species and habitats. Species dependence on a habitat is indicated in the cells under each habitat; if there is no entry for a species stage, then it is not dependent on that habitat. The table uses color to define a range across low (green), moderate (yellow), high (orange), and very high (red) climate vulnerability for species (first column) and habitats (all other columns). This color scheme also denotes the level of species dependence on a habitat.

```{r class.source="risktable"}
#Read in Mid-Atlantic table
mid <- as.data.table(readxl::read_xlsx(here('data-raw/habitat_vulnerability_2022.xlsx'),
                         sheet = 'Mid-Atlantic', skip = 1))

#Identify individual species climate vulnerability
vhigh.vul <- unique(mid[`Species Vulnerability Rank (FCVA)` == 'Very high', Species])
high.vul  <- unique(mid[`Species Vulnerability Rank (FCVA)` == 'High', Species])
mod.vul   <- unique(mid[`Species Vulnerability Rank (FCVA)` == 'Moderate', Species])
low.vul   <- unique(mid[`Species Vulnerability Rank (FCVA)` == 'Low', Species])

#Grab habitat vulnerability
hab.vul <- unique(mid[, c('Habitat Name', 'Habitat Vulnerability Rank (HVCA)')])
habitats <- hab.vul[, 'Habitat Name']
hab.vul <- data.table('Habitat Vulnerability' = c(NA, NA, t(hab.vul[, 'Habitat Vulnerability Rank (HVCA)'])))

#Rearrange table
mid <- data.table::melt.data.table(mid, id.vars = c('Habitat Name', 'Species'),
                                   measure.vars = c('Eggs/Larva', 'Juvenile/YOY',
                                                    'Adult', 'Spawning Adult'),
                                   variable.name = 'Stage', value.name = 'Dependence')
mid[, Habitat := as.factor(mid$'Habitat Name')]
mid <- data.table::dcast.data.table(mid, Species + Stage ~ Habitat,
                                    value.var = 'Dependence')
setcolorder(mid, c('Species', 'Stage', habitats$'Habitat Name'))

#Add Habitat Vulnerbaility
hab.table <- rbindlist(list(as.data.table(t(hab.vul)), mid), use.names = F)
#Add names back in
names(hab.table) <- names(mid)
hab.table <- hab.table %>% mutate_if(is.character, str_trim) #one cell has a space, not getting color


#generalize to all
# hab.table %>%
#   mutate_at(vars(Species), function(x){ #Color code species based on climate vul
#     case_when(x %in% low.vul ~ cell_spec(x, format = "html", color = "black",
#                                      background = "lightgreen"),
#               x %in% mod.vul ~ cell_spec(x, format = "html", color = "black",
#                                      background = "yellow"),
#               x %in% high.vul ~ cell_spec(x, format = "html", color = "black",
#                                      background = "orange"),
#               x %in% vhigh.vul ~ cell_spec(x, format = "html", color = "white",
#                                      background = "red"))}) %>%
#   mutate_at(vars(-c(Species, Stage)), function(x){ #Color code base on dependence
#     case_when(x == "Low" ~ cell_spec(x, format = "html", color = "black",
#                                      background = "lightgreen"),
#               x == "Moderate" ~ cell_spec(x, format = "html", color = "black",
#                                      background = "yellow"),
#               x == "High" ~ cell_spec(x, format = "html", color = "black",
#                                      background = "orange"),
#               x == "Very high" ~ cell_spec(x, format = "html", color = "white",
#                                      background = "red"))}) %>%
#   kable(format = "html", escape = F, table.attr='class="risktable"') %>%
#   kable_styling(font_size = 14, fixed_thead = T)

```

```{r midDT}
#from examples here https://rstudio.github.io/DT/010-style.html

#color coding cells works
#need to add column with species vulnerability and color code species by it but not show it, like #hideV6 example at link
#if we show it we can sort by species vulnerability though... alphabetical doesnt work so use numbers to go lowest-highest
#still need to color code headers--done with spans below
hab.table <- hab.table %>%
  filter(Species != "") %>% #remove row with habitat vul because headers now in color
  mutate(spVul = #Color code species based on climate vul
    # case_when(Species %in% low.vul ~ "low.vul",
    #           Species %in% mod.vul ~ "mod.vul",
    #           Species %in% high.vul ~ "high.vul",
    #           Species %in% vhigh.vul ~ "vhigh.vul")
    # )
    case_when(Species %in% low.vul ~ 1,
              Species %in% mod.vul ~ 2,
              Species %in% high.vul ~ 3,
              Species %in% vhigh.vul ~ 4)
    )

colnames(hab.table)[c(3:4)] <- paste0('<span style="box-shadow: -10px 0 0 red, 10px 0 0 red; background-color:red; color:white">',colnames(hab.table)[c(3:4)],'</span>')
colnames(hab.table)[c(5:11)] <- paste0('<span style="box-shadow: -10px 0 0 orange, 10px 0 0 orange; background-color:',c("orange"),'">',colnames(hab.table)[c(5:11)],'</span>')
colnames(hab.table)[c(12:16)] <- paste0('<span style="box-shadow: -10px 0 0 yellow, 10px 0 0 yellow; background-color:',c("yellow"),'">',colnames(hab.table)[c(12:16)],'</span>')

dt <- datatable(hab.table, rownames = FALSE, escape = FALSE,
          extensions = c('Scroller', 'FixedColumns'), #add scrolling and fixed column
          options = list(deferRender = TRUE,
          scrollY = 300, scroller = TRUE,
          dom = 't',scrollX = TRUE,fixedColumns = TRUE)
          #, options = list(columnDefs = list(list(targets = 16, visible = FALSE)))
  ) %>% 
  formatStyle(
    names(hab.table)[3:4],
    backgroundColor = '#FF000080'
  ) %>%
  formatStyle(
    names(hab.table)[5:11],
    backgroundColor = '#FFA50080'
  ) %>%
  formatStyle(
    names(hab.table)[12:16],
    backgroundColor = '#FFFF0080'
  ) %>%

  formatStyle(#Color code base on dependence
    'Species', "spVul",
    backgroundColor = styleEqual(#c("low.vul", "mod.vul", "high.vul", "vhigh.vul"),
                                 c(1,2,3,4),
                                 c('lightgreen', 'yellow', 'orange', 'red')), #Color code base on dependence)
    color = styleEqual(c(1,2,3,4),
                                 c('black', 'black', 'black', 'white'))
    ) %>%
    formatStyle(
    names(hab.table),
    backgroundColor = styleEqual(c("Low", "Moderate", "High", "Very high"), 
                                 c('lightgreen', 'yellow', 'orange', 'red')), 
    color = styleEqual(c("Low", "Moderate", "High", "Very high"), 
                                 c('black', 'black', 'black', 'white'))
  )

dt
```

#### New England Habitat Vulnerability Table



```{r class.source="risktable"}
#Read in New England table
ne <- as.data.table(readxl::read_xlsx(here('data-raw/habitat_vulnerability.xlsx'),
                         sheet = 'New England', skip = 1))

#Identify individual species climate vulnerability
vhigh.vul <- unique(ne[`Species Vulnerability Rank (FCVA)` == 'Very high', Species])
high.vul  <- unique(ne[`Species Vulnerability Rank (FCVA)` == 'High', Species])
mod.vul   <- unique(ne[`Species Vulnerability Rank (FCVA)` == 'Moderate', Species])
low.vul   <- unique(ne[`Species Vulnerability Rank (FCVA)` == 'Low', Species])

#Grab habitat vulnerability
hab.vul <- unique(ne[, c('Habitat Name', 'Habitat Vulnerability Rank (HVCA)')])
habitats <- hab.vul[, 'Habitat Name']
hab.vul <- data.table('Habitat Vulnerability' = c(NA, NA, t(hab.vul[, 'Habitat Vulnerability Rank (HVCA)'])))

#Rearrange table
ne <- data.table::melt.data.table(ne, id.vars = c('Habitat Name', 'Species'),
                                   measure.vars = c('Eggs/Larva', 'Juvenile/YOY',
                                                    'Adult', 'Spawning Adult'),
                                   variable.name = 'Stage', value.name = 'Dependence')
ne[, Habitat := as.factor(ne$'Habitat Name')]
ne <- data.table::dcast.data.table(ne, Species + Stage ~ Habitat,
                                    value.var = 'Dependence')
setcolorder(ne, c('Species', 'Stage', habitats$'Habitat Name'))

#Add Habitat Vulnerbaility
hab.table <- rbindlist(list(as.data.table(t(hab.vul)), ne), use.names = F)
#Add names back in
names(hab.table) <- names(ne)
hab.table <- hab.table %>% mutate_if(is.character, str_trim) #one cell has a space, not getting color

#generalize to all
# hab.table %>%
#   mutate_at(vars(Species), function(x){ #Color code species based on climate vul
#     case_when(x %in% low.vul ~ cell_spec(x, format = "html", color = "black",
#                                      background = "lightgreen"),
#               x %in% mod.vul ~ cell_spec(x, format = "html", color = "black",
#                                      background = "yellow"),
#               x %in% high.vul ~ cell_spec(x, format = "html", color = "black",
#                                      background = "orange"),
#               x %in% vhigh.vul ~ cell_spec(x, format = "html", color = "white",
#                                      background = "red"))}) %>%
#   mutate_at(vars(-c(Species, Stage)), function(x){ #Color code base on dependence
#     case_when(x == "Low" ~ cell_spec(x, format = "html", color = "black",
#                                      background = "lightgreen"),
#               x == "Moderate" ~ cell_spec(x, format = "html", color = "black",
#                                      background = "yellow"),
#               x == "High" ~ cell_spec(x, format = "html", color = "black",
#                                      background = "orange"),
#               x == "Very high" ~ cell_spec(x, format = "html", color = "white",
#                                      background = "red"))}) %>%
#   kable(format = "html", escape = F, table.attr='class="risktable"') %>%
#   kable_styling(font_size = 14, fixed_thead = T)

```

```{r neDT}

#color coding cells works
#need to add column with species vulnerability and color code species by it but not show it, like #hideV6 example at link
#if we show it we can sort by species vulnerability though... alphabetical doesnt work so use numbers to go lowest-highest
hab.table <- hab.table %>%
  mutate(spVul = #Color code species based on climate vul
    # case_when(Species %in% low.vul ~ "low.vul",
    #           Species %in% mod.vul ~ "mod.vul",
    #           Species %in% high.vul ~ "high.vul",
    #           Species %in% vhigh.vul ~ "vhigh.vul")
    # )
    case_when(Species %in% low.vul ~ 1,
              Species %in% mod.vul ~ 2,
              Species %in% high.vul ~ 3,
              Species %in% vhigh.vul ~ 4)
    )


datatable(hab.table, rownames = FALSE, 
           extensions = c('Scroller', 'FixedColumns'), #add scrolling and fixed column
          options = list(deferRender = TRUE,
          scrollY = 200, scroller = TRUE,
          dom = 't',scrollX = TRUE,fixedColumns = TRUE)
          #, options = list(columnDefs = list(list(targets = 16, visible = FALSE)))
  ) %>% 
  formatStyle(#Color code base on dependence
    'Species', "spVul",
    backgroundColor = styleEqual(#c("low.vul", "mod.vul", "high.vul", "vhigh.vul"),
                                 c(1,2,3,4),
                                 c('lightgreen', 'yellow', 'orange', 'red')) #Color code base on dependence)
    ) %>%
    formatStyle(
    names(hab.table)[2:15],
    backgroundColor = styleEqual(c("Low", "Moderate", "High", "Very high"), 
                                 c('lightgreen', 'yellow', 'orange', 'red'))
  )


```

#### Data Sources

Farr, E. R., Johnson, M. R., Nelson, M. W., Hare, J. A., Morrison, W. E., Lettrich, M. D., Vogt, B., et al. 2021. An assessment of marine, estuarine, and riverine habitat vulnerability to climate change in the Northeast U.S. PLOS ONE, 16: e0260654. Public Library of Science.

Hare, J. A., Morrison, W. E., Nelson, M. W., Stachura, M. M., Teeters, E. J., Griffis, R. B., Alexander, M. A., et al. 2016. A Vulnerability Assessment of Fish and Invertebrates to Climate Change on the Northeast U.S. Continental Shelf. PLOS ONE, 11: e0146756.

Kritzer, J. P., DeLucia, M.-B., Greene, E., Shumway, C., Topolski, M. F., Thomas-Blate, J., Chiarella, L. A., et al. 2016. The Importance of Benthic Habitats for Coastal Fisheries. BioScience, 66: 274–284.
