---
title: "Environmental and LTL Indicators"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    theme: lumen
---

```{r setup, include=FALSE}

#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center') #allows for inserting R code into captions

#Plotting and data libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(ecodata)
library(here)
library(kableExtra)
library(ggrepel)
library(stringr)
library(patchwork)
library(grid)

#GIS libraries
library(sf)
library(rgdal)
library(raster)
library(rnaturalearth)


#General inline text input for report

#Council
council <- "Mid-Atlantic Fishery Management Council"
council_abbr <- "MAFMC"

#Region identifiers
epu <- "Mid-Atlantic Bight"
epu_abbr <- "MAB"
region <- "Mid-Atlantic"
region_abbr <- "MA" #Some commercial data organized by "MA" or "NE" regions, not by EPU 


```

```{r GIS-setup}

#CRS
crs <- "+proj=longlat +lat_1=35 +lat_2=45 +lat_0=40 +lon_0=-77 +x_0=0 +y_0=0 +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"

#Coastline shapefile
coast <- ne_countries(scale = 10,
                          continent = "North America",
                          returnclass = "sf") %>%
             sf::st_transform(crs = crs)

#State polygons
ne_states <- ne_states(country = "united states of america",
                                      returnclass = "sf") %>%
  sf::st_transform(crs = crs)

#EPU shapefile
epu_sf <- ecodata::epu_sf %>% 
  filter(EPU %in% c("MAB","GB","GOM"))

#Map line parameters
map.lwd <- 0.4

# Set lat/lon window for maps
xmin = -77
xmax = -65
ymin = 36
ymax = 45
xlims <- c(xmin, xmax)
ylims <- c(ymin, ymax)

#Time series constants
shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
feeding.guilds <- c("Apex Predator","Piscivore","Planktivore","Benthivore","Benthos")
x.shade.min <- 2009
x.shade.max <- 2018
#Function for custom ggplot facet labels
label <- function(variable,value){
  return(facet_names[value])
}
```

Trend lines are shown when slope is significantly different from 0 at the p < 0.05 level. An orange line signifies an overall positive trend, and purple signifies a negative trend. Note that in the final report we will only test for trend when N >= 30. However, I have relaxed that requirement for the purposes of this document so that trends are highlighted when N >= 20. **This means that some trends shown here will not be present in the final document**. Dashed lines represent mean values of time series unless the indicator is an anomaly, in which case the dashed line is equal to 0. Shaded regions indicate the past ten years. If there are no new data for 2018, the shaded region will still cover this time period.

## Mid-Atlantic Bight

### Surface Winds {.tabset .tabset-fade}

#### Total wind speed

```{r TWS, echo = F}

data.dir <- here::here("data")
load(file.path(data.dir,"wind.rda"))

all_winds <- wind %>% 
  filter(str_detect(Var, "total wind speed|hcly|tke"),
         EPU == epu_abbr) %>% 
  group_by(Var) %>% 
  mutate(hline = mean(Value, na.rm = T)) %>%
  mutate(Value = Value - hline) %>% 
  mutate(hline = 0) %>% 
   ungroup() %>% 
  mutate(Season = str_to_title(word(Var, -1))) 
 
all_winds$Season <- factor(all_winds$Season, levels = c("Winter",
                                                        "Spring",
                                                        "Summer",
                                                        "Fall"))
(tws <- all_winds %>%
  filter(str_detect(Var, "total wind speed")) %>% #filter
  ggplot()+ #plot
      #Highlight last ten years
    
  annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab("Wind speed anomaly (m/s)") +
  ggtitle("Total Wind Speed") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    scale_y_continuous(limits = c(-2,2))+
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2)+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12)))
```

#### Helicity ("storminess")

```{r Helicity, echo = F}
(helicity <- all_winds %>%
  filter(str_detect(Var, "hcly")) %>% #filter
  ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab(expression("Relative Helicity (m"^2*" sec"^-2*")")) +
  ggtitle("Storminess") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12)))
```

#### Turbulent kinetic energy

```{r TKE, echo  = F}
(tke <- all_winds %>%
  filter(str_detect(Var, "tke")) %>% #filter
  ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab(expression("Kinetic Energy (J kg"^-1*")")) +
  ggtitle("Turbulent Kinetic Energy") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12)))
```

### Ocean Temperature {.tabset .tabset-fade}

#### SST

```{r MAB-SST-insitu, fig.width = 8, fig.height = 7}
annotation_custom2 <- function (grob, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, data) 
{
  layer(data = data, stat = StatIdentity, position = PositionIdentity, 
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE, params = list(grob = grob, 
                                          xmin = xmin, xmax = xmax, 
                                          ymin = ymin, ymax = ymax))
}

#EPU shapefile
epu_sf <- ecodata::epu_sf %>% 
  filter(EPU %in% c("MAB"))

#Map line parameters
map.lwd <- 0.4

# Set lat/lon window for maps
xmin = -81
xmax = -66
ymin = 35.5
ymax = 43
xlims <- c(xmin, xmax)
ylims <- c(ymin, ymax)
sst <- seasonal_oisst_anom_gridded 

sst$Season <- factor(sst$Season, levels = c("Winter",
                                            "Spring",
                                            "Summer",
                                            "Fall"))
sst_map <- 
  ggplot() +
  geom_tile(data = sst, aes(x = Longitude, y = Latitude,fill = Value)) +
  geom_sf(data = coast, size = map.lwd) +
  geom_sf(data = epu_sf, fill = "transparent", size = map.lwd) +
  scale_fill_gradient2(name = "Temp.\nAnomaly (°C)",
                       low = scales::muted("blue"),
                       mid = "white",
                       high = scales::muted("red"),
                       limits = c(-5,5)) +
  coord_sf(crs = crs, xlim = xlims, ylim = ylims) +
  facet_wrap(Season~.) +
  theme_map() +
  ggtitle("SST anomaly (2018)") +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=0.75),
        legend.key = element_blank(),
        axis.title = element_text(size = 11),
        strip.background = element_blank(),
        strip.text=element_text(hjust=0),
        axis.text = element_text(size = 8))

winter_anom <-  ggplotGrob( seasonal_oisst_anom %>% 
                              filter(EPU == "MAB",
                                     str_detect(Var, "winter")) %>% 
                              mutate(hline = mean(Value)) %>% 
                              ggplot(aes(x = Time, y = Value)) +
                              annotate("rect", fill = shade.fill, alpha = shade.alpha,
                                       xmin = x.shade.min , xmax = x.shade.max,
                                       ymin = -Inf, ymax = Inf) +
                              geom_line() +
                              geom_point() +
                              geom_gls(alpha = trend.alpha + 0.25) +
                              ylab("SST anomaly (°C)")+
                              scale_x_continuous(expand = c(0.01, 0.01)) +
                              geom_hline(aes(yintercept = hline)) +
                              theme_ts()+
                              theme(axis.title = element_text(size = 6),
                                    axis.text = element_text(size = 6),
                                    panel.background = element_rect(fill = "transparent"), 
                                    plot.background = element_rect(fill = "transparent", color = NA), 
                                    panel.grid.major = element_blank(),
                                    panel.grid.minor = element_blank(), 
                                    legend.background = element_rect(fill = "transparent"), 
                                    legend.box.background = element_rect(fill = "transparent"), 
                                    legend.key = element_rect(fill = "transparent", colour = NA), 
                                    axis.line = element_blank(),
                                    panel.border = element_blank())
)

spring_anom <-  ggplotGrob( seasonal_oisst_anom %>% 
                              filter(EPU == "MAB",
                                     str_detect(Var, "spring")) %>% 
                              mutate(hline = mean(Value)) %>% 
                              ggplot(aes(x = Time, y = Value)) +
                              annotate("rect", fill = shade.fill, alpha = shade.alpha,
                                       xmin = x.shade.min , xmax = x.shade.max,
                                       ymin = -Inf, ymax = Inf) +
                              geom_line() +
                              geom_point() +
                              geom_gls(alpha = trend.alpha + 0.25) +
                              ylab("SST anomaly (°C)")+
                              scale_x_continuous(expand = c(0.01, 0.01)) +
                              geom_hline(aes(yintercept = hline)) +
                              theme_ts()+
                              theme(axis.title = element_text(size = 6),
                                    axis.text = element_text(size = 6),
                                    panel.background = element_rect(fill = "transparent"), 
                                    plot.background = element_rect(fill = "transparent", color = NA),
                                    panel.grid.major = element_blank(),
                                    panel.grid.minor = element_blank(), 
                                    legend.background = element_rect(fill = "transparent"), 
                                    legend.box.background = element_rect(fill = "transparent"), 
                                    legend.key = element_rect(fill = "transparent", colour = NA), 
                                    axis.line = element_blank(),
                                    panel.border = element_blank())
)

summer_anom <-  ggplotGrob( seasonal_oisst_anom %>% 
                              filter(EPU == "MAB",
                                     str_detect(Var, "summer")) %>% 
                              mutate(hline = mean(Value)) %>% 
                              ggplot(aes(x = Time, y = Value)) +
                              annotate("rect", fill = shade.fill, alpha = shade.alpha,
                                       xmin = x.shade.min , xmax = x.shade.max,
                                       ymin = -Inf, ymax = Inf) +
                              geom_line() +
                              geom_point() +
                              geom_gls(alpha = trend.alpha + 0.25) +
                              ylab("SST anomaly (°C)")+
                              scale_x_continuous(expand = c(0.01, 0.01)) +
                              geom_hline(aes(yintercept = hline)) +
                              theme_ts()+
                              theme(axis.title = element_text(size = 6),
                                    axis.text = element_text(size = 6),
                                    panel.background = element_rect(fill = "transparent"),
                                    plot.background = element_rect(fill = "transparent", color = NA),
                                    panel.grid.major = element_blank(), 
                                    panel.grid.minor = element_blank(), 
                                    legend.background = element_rect(fill = "transparent"),
                                    legend.box.background = element_rect(fill = "transparent"), 
                                    legend.key = element_rect(fill = "transparent", colour = NA), 
                                    axis.line = element_blank(),
                                    panel.border = element_blank())
)

fall_anom <-  ggplotGrob( seasonal_oisst_anom %>% 
                            filter(EPU == "MAB",
                                   str_detect(Var, "fall")) %>% 
                            mutate(hline = mean(Value)) %>% 
                            ggplot(aes(x = Time, y = Value)) +
                            annotate("rect", fill = shade.fill, alpha = shade.alpha,
                                     xmin = x.shade.min , xmax = x.shade.max,
                                     ymin = -Inf, ymax = Inf) +
                            geom_line() +
                            geom_point() +
                            geom_gls(alpha = trend.alpha + 0.25) +
                            ylab("SST anomaly (°C)")+
                            scale_x_continuous(expand = c(0.01, 0.01)) +
                            geom_hline(aes(yintercept = hline)) +
                            theme_ts()+
                            theme(axis.title = element_text(size = 6),
                                  axis.text = element_text(size = 6),
                                  panel.background = element_rect(fill = "transparent"), 
                                  plot.background = element_rect(fill = "transparent", color = NA), 
                                  panel.grid.major = element_blank(), 
                                  panel.grid.minor = element_blank(), 
                                  legend.background = element_rect(fill = "transparent"), 
                                  legend.box.background = element_rect(fill = "transparent"),
                                  legend.key = element_rect(fill = "transparent", colour = NA),
                                  axis.line = element_blank(),
                                  panel.border = element_blank())
)

sst_map + 
  annotation_custom2(grob = winter_anom,  xmin=-82, xmax=-73.75,
                     ymin=39, ymax=43.5, data = data.frame(Season = "Winter")) +
  annotation_custom2(grob = spring_anom,  xmin=-82, xmax=-73.75,
                     ymin=39, ymax=43.5, data = data.frame(Season = "Spring")) +
  annotation_custom2(grob = summer_anom,  xmin=-82, xmax=-73.75,
                     ymin=39, ymax=43.5, data = data.frame(Season = "Summer")) +
  annotation_custom2(grob = fall_anom,  xmin=-82, xmax=-73.75,
                     ymin=39, ymax=43.5, data = data.frame(Season = "Fall"))

```

#### Bottom temperature

```{r MAB-bot-temp, fig.width = 6, fig.asp = .55}
temp_anom <- oceantemp_insitu %>% 
  filter(EPU == epu_abbr) %>% 
  complete(Time = full_seq(min(oceantemp_insitu$Time):max(oceantemp_insitu$Time),1),
           nesting(Var)) %>% 
  mutate(hline = 0)

bot_temp_insitu <- temp_anom %>%
 filter(Var == "bottom temp anomaly in situ") %>%
ggplot2::ggplot() +
  annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Temperature (°C)") +
  ggtitle("Bottom temp. anomaly") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))+
  geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
 theme_ts()


bot_temp_insitu 

```

### Ocean Salinity & Stratification {.tabset .tabset-fade}

#### Surface salinity

```{r MAB-surfsal, fig.width = 8, fig.asp = .35}
sal_anom <- oceansal_insitu %>% 
  filter(EPU == epu_abbr) %>% 
  complete(Time = full_seq(min(oceansal_insitu$Time):max(oceansal_insitu$Time),1),
           nesting(Var)) %>% 
  mutate(hline = 0)

surf_sal_insitu <- sal_anom %>%
  filter(Var == "surface salinity anomaly in situ") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Salinity (PSU)") +
  ggtitle("Surface salinity anomaly (in situ)") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

sal_oi <- oceansal_oi %>% 
  filter(EPU == epu_abbr) %>%
  group_by(Var) %>% 
  mutate(hline = mean(Value))

surf_sal_oi <- sal_oi %>%
  filter(Var == "surface sal OI") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  #geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab(expression("")) +
  ggtitle("Surface salinity (optimally interpolated)") +
  scale_x_continuous(expand = c(0.01, 0.01), limits = c(1977,2018)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

surf_sal_insitu + surf_sal_oi +
  plot_layout(ncol =  2) &
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

```

#### Bottom salinity

```{r MAB-botsal, fig.width = 8, fig.asp = .35}

bot_sal_insitu <- sal_anom %>%
  filter(Var == "bottom salinity anomaly in situ") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Salinity (PSU)") +
  ggtitle("Bot. salinity anomaly (in situ)") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

bot_sal_oi <- sal_oi %>%
  filter(Var == "bottom sal OI") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
 # geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab(expression("")) +
  ggtitle("Bot. salinity (optimally interpolated)") +
  scale_x_continuous(expand = c(0.01, 0.01), limits = c(1977,2018)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

bot_sal_insitu + bot_sal_oi +
  plot_layout(ncol =  2) &
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

```

#### Stratification

```{r MAB-strat}
strat <- ecodata::stratification %>% 
  filter(EPU == epu_abbr) %>% 
  mutate(hline = mean(Value, na.rm = T))

strat %>% 
  ggplot(aes(x = Time, y = Value)) +
       annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line() +
  geom_point() +
 # geom_gls() +
    ylab(expression("Stratification (kg m"^-3*")")) +
  ggtitle("Stratification") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() 

```


### Chlorophyll and Primary Productivity {.tabset .tabset-fade}

#### Primary production

```{r PP-OCCI, fig.width=8}
out_pp <- ecodata::chl_pp %>% 
  filter(EPU == epu_abbr,
         str_detect(Var, "MONTHLY_PPD_MEDIAN")) %>% 
  separate(.,Time, into = c("Year","Month"), sep = 4) %>% 
    mutate(Month = plyr::mapvalues(Month, from = c("01","02","03","04","05","06",
                                                   "07","08","09","10","11","12"),
                                   to = c(month.abb))) %>% 
  group_by(Month) %>% 
  mutate(hline = mean(Value))

out_pp$Month <- factor(out_pp$Month, levels = month.abb)


(pp_cci <- ggplot(out_pp) +
   # geom_gls(aes(x = Year, y = Value, group = Month))+
    geom_point(aes(x = Year, y = Value, group = Month)) +
    geom_line(aes(x = Year, y = Value, group = Month)) +
    scale_x_discrete(name = "Time", breaks = seq(min(out_pp$Year),max(out_pp$Year),10)) +  
        geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
    facet_wrap(Month~., ncol = 6) +
    ggtitle("Monthly median PPD") +
    ylab(expression("PP (gC m"^-2*" d"^-1*")")) +
    theme_facet() +
    theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm")))
```


```{r chl-monthly, eval = F}
out_chl <- ecodata::chl_pp %>% 
  filter(EPU == epu_abbr,
         str_detect(Var, "MONTHLY_CHLOR_A_MEDIAN")) %>% 
  separate(.,Time, into = c("Year","Month"), sep = 4) %>% 
    mutate(Month = plyr::mapvalues(Month, from = c("01","02","03","04","05","06",
                                                   "07","08","09","10","11","12"),
                                   to = c(month.abb)))%>% 
  group_by(Month) %>% 
  mutate(hline = mean(Value))
out_chl$Month <- factor(out_chl$Month, levels = month.abb)


(chl_cci <- ggplot(out_chl) +
   # geom_gls(aes(x = Year, y = Value, group = Month))+
    geom_point(aes(x = Year, y = Value, group = Month)) +
    geom_line(aes(x = Year, y = Value, group = Month)) +
    scale_x_discrete(name = "Time", breaks = seq(min(out_chl$Year),max(out_chl$Year),10)) +  
            geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
    facet_wrap(Month~., ncol = 6) +
    ggtitle("Monthly median CHL") +
    ylab(expression("CHL (mg m"^-3*")")) +
    theme_facet() +
    theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm")))

```


#### Sesonal chlorophyll *a* & primary production

```{r mab-chl-weekly,fig.cap = "Weekly chlorophyll concentrations and primary productivity in the Mid-Atlantic are shown for by the colored line for 2018. The long-term mean is shown in black and shading indicates +/- 1 sample SD."}

interp_chl_pp <- function(epu, int_step, year = 2018, Variable){
  out <- ecodata::chl_pp %>% 
    filter(str_detect(Var,Variable),
           EPU == epu) %>% 
    separate(.,Time, c("Year","Week"),sep = 4) %>% 
    filter(Year == year) %>% 
    group_by(EPU) %>% 
    mutate(Week = 1:length(Year))  %>% 
    complete(Week = full_seq(1:52,1),
             nesting(Var))
  
  out_int <- data.frame(Value = approx(out$Week,
                                       out$Value,
                                       seq(1,max(out$Week),
                                           length.out = length(out$Week)*int_step))$y,
                        Time = seq(1,max(out$Week),length.out = length(out$Week)*int_step),
                        Week = rep(1:52,each = int_step))
  
  ltm_out <- ecodata::chl_pp %>% 
    filter(str_detect(Var,Variable),
           EPU == epu) %>% 
    separate(.,Time, c("Year","Week"),sep = 4) %>% 
    group_by(Week) %>% 
    dplyr::summarise(LTM = mean(Value, na.rm = T),
                     SD = sd(Value, na.rm = T)) %>% 
    mutate(Week = 1:length(Week),
           sd.low = LTM - SD,
           sd.high = LTM + SD) 
  
  ltm_out_int <- data.frame(LTM = approx(ltm_out$Week,
                                         ltm_out$LTM,
                                         seq(1,max(ltm_out$Week),
                                             length.out = length(ltm_out$Week)*int_step))$y,
                            sd.high = approx(ltm_out$Week,
                                             ltm_out$sd.high,
                                             seq(1,max(ltm_out$Week),
                                                 length.out = length(ltm_out$Week)*int_step))$y,
                            sd.low = approx(ltm_out$Week,
                                            ltm_out$sd.low,
                                            seq(1,max(ltm_out$Week), 
                                                length.out = length(ltm_out$Week)*int_step))$y,
                            Time = seq(1,max(out$Week),length.out = length(out$Week)*int_step),
                            Week = rep(1:52,each = int_step)) %>%
    left_join(.,out_int, by = c("Week","Time")) %>% 
    mutate(status = ifelse(Value < sd.high & Value > sd.low, "near_mean",
                           ifelse(Value > sd.high, "high",
                                  ifelse(Value < sd.low,"low",NA))),
           group = "PLOT")
  
  return(ltm_out_int)
}


MAB_chl <- interp_chl_pp(epu = "MAB", int_step = 1000,Variable = "WEEKLY_CHLOR_A_MEDIAN MODIS-Aqua PAN")

MAB_chl_weekly <- ggplot(data = MAB_chl) +
  geom_line(aes(x = Time, y = LTM)) +
  geom_ribbon(aes(x = Time, ymin = sd.low, ymax = sd.high), 
              alpha = 0.1,
              fill = "grey1") +
  geom_line(aes(x = Time, y = Value, color = status, group = group),
            size = 1) +
  ggtitle(expression("Chlorophyll"~italic(a)~"")) +
  guides(color = F) +
  xlab("")+
  ylab(expression("CHL (mg m"^-3*")")) +
  scale_x_continuous(breaks = seq(0,50,10),
                   labels = c("January","March","May","July","October","December")) +
  scale_color_manual(values = c("#ef8a62","#2c7fb8","#a1d99b"))+
  theme_ts()

MAB_pp <- interp_chl_pp(epu = "MAB", int_step = 1000,Variable =  "WEEKLY_PPD_MEDIAN")

MAB_pp_weekly <- ggplot(data = MAB_pp) +
  geom_line(aes(x = Time, y = LTM)) +
  geom_ribbon(aes(x = Time, ymin = sd.low, ymax = sd.high),
              alpha = 0.1,
              fill = "grey1") +
  geom_line(aes(x = Time, y = Value, color = status, group = group),
            size = 1) +
  ggtitle(expression("Primary production")) +
  guides(color = F) +
  xlab("")+
  ylab(expression("PP (gC m"^-2*" d"^-1*")")) +
  scale_x_continuous(breaks = seq(0,50,10),
                   labels = c("January","March","May","July","October","December")) +
  scale_color_manual(values = c("#ef8a62","#2c7fb8","#a1d99b"))+
  theme_ts()

MAB_chl_weekly + MAB_pp_weekly + plot_layout(ncol = 1)

```



### Chesapeake Bay Water Quality Attainment

```{r cb-attainment,fig.width = 6, fig.asp = 0.55, fig.cap = "Water quality attainment in Chesapeake Bay following rolling three year assessment periods."}

minlab <- seq(1985,2015,5)
maxlab <- seq(1987,2017,5)



ches_bay_wq %>% 
  mutate(hline = mean(Value)) %>% 
  ggplot(aes(x = Time, y = Value)) +
       annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line() +
  geom_point() +
  geom_gls() +
    ylab(expression("Estimated attainment, percent")) +
  ggtitle("Chesapeake Bay Water Quality Attainment") +
  scale_x_continuous(breaks = minlab,labels = paste0(minlab,"-",maxlab),expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts()
```


### Zooplankton {.tabset .tabset-fade}

#### Abundance anomaly

```{r MAB-zoo-abund, fig.width = 8, fig.asp = .35}
zoo_abund <- ecodata::zoo_anom_sli %>% 
  filter(EPU == epu_abbr,
         !str_detect(Var, "small-large")) %>% 
  mutate(hline = 0,
         Var = str_to_title(str_remove(Var, "anomaly"))) 

zoo_abund %>% 
  ggplot(aes(x = Time, y = Value, group = Var)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_gls() +
  geom_line() +
  geom_point() +
  ylab("Abundance anomaly") +
  ggtitle("Zooplankton abundance anomaly") +
  facet_wrap(Var~., ncol = 3) +
  scale_x_continuous(expand = c(0.01, 0.01))+
      geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_facet() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))
  

```

#### Small-large index

```{r MAB-sli}
sli <- ecodata::zoo_anom_sli %>% 
  filter(EPU == epu_abbr,
         str_detect(Var, "small-large")) %>% 
  mutate(hline = 0) 

pp_anom <- ecodata::chl_pp %>% 
  filter(str_detect(Var, "ANNUAL_PPD_RATIO_ANOMALY"),
         EPU == "MAB") %>% 
  mutate(hline = 1,
         Time = as.numeric(as.character(Time)),
         Var = "ANNUAL_PPD_RATIO_ANOMALY")

sli_plt <- sli %>% 
  ggplot(aes(x = Time, y = Value, group = Var)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line() +
  geom_point() +
  guides(color = F) +
  xlab("")+
  ylab("Small-large abundance") +
  ggtitle("Small-large copepod abundance") +
    scale_x_continuous(expand = c(0.01, 0.01), limits = c(1998, 2018))+
      geom_hline(aes(yintercept = hline,
                     group = Var),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_ts() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))

pp_anom_plt <- pp_anom %>% 
    ggplot(aes(x = Time, y = Value, group = Var)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line() +
  geom_point() +
  guides(color = F) +
  ylab("Anomaly ratio") +
  ggtitle("Primary production anomaly ratio") +
    scale_x_continuous(expand = c(0.01, 0.01), limits = c(1998, 2018))+
    scale_y_continuous(limits = c(0.8,1.2)) +
      geom_hline(aes(yintercept = hline,
                     group = Var),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_ts() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))


sli_plt + pp_anom_plt + plot_layout(ncol = 1) & theme(plot.margin = margin(t = 0))
```

#### OI abundance

```{r MAB-oi-zoo, fig.width=8}
facet_names <- list(
  'centropages spring'=expression(paste(italic("Centropages "), "spring")),
  'centropages fall'=expression(paste(italic("Centropages "), "fall")),
  'temora spring'=expression(paste(italic("Temora "), "spring")),
  'temora fall' = expression(paste(italic("Temora "), "fall")),
  'pseudocalanus spring'=expression(paste(italic("Pseudocalanus "), "spring")),
  'pseudocalanus fall' = expression(paste(italic("Pseudocalanus "), "fall")))

zoo_oi_mab <- ecodata::zoo_oi %>% 
  filter(!str_detect(Var,"SD"),
         EPU == "MAB") %>% 
  mutate(Var = str_remove(Var, " zoo")) %>% 
  group_by(EPU, Var) %>% 
  mutate(hline = mean(Value, na.rm = T))

zoo_oi_mab %>%  
ggplot(aes(x = Time, y = Value, group = Var)) +
  geom_gls() +
  geom_line()+
  geom_point()+
  ylab(expression("Abundance log num m"^-3*"")) +
  facet_wrap(Var ~ ., ncol = 2,scales = "free_y", labeller = label) +
  ggtitle("Zooplankton abundance (OI)") +
  theme_facet() +
    theme(strip.text=element_text(hjust=0,
                                face = "italic"))

```

## Gulf of Maine & Georges Bank

### Surface Winds {.tabset .tabset-fade}

#### Total wind speed

```{r NE-TWS, echo = F, fig.width=8, fig.height=8}

all_winds <- wind %>% 
  filter(str_detect(Var, "total wind speed|hcly|tke"),
         EPU %in% c("GOM","GB")) %>% 
  group_by(Var) %>% 
  mutate(hline = mean(Value, na.rm = T)) %>%
   ungroup() %>% 
  mutate(Season = str_to_title(word(Var, -1))) 
 
all_winds$Season <- factor(all_winds$Season, levels = c("Winter",
                                                        "Spring",
                                                        "Summer",
                                                        "Fall"))
tws_gom <- all_winds %>%
  filter(EPU == "GOM",
         str_detect(Var, "total wind speed")) %>% #filter
  ggplot()+ #plot
      #Highlight last ten years
    
  annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab("Wind Speed (m/s)") +
  ggtitle("GOM Total Wind Speed") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

tws_gb <- all_winds %>%
  filter(EPU == "GB",
         str_detect(Var, "total wind speed")) %>% #filter
  ggplot()+ #plot
      #Highlight last ten years
    
  annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab("Wind Speed (m/s)") +
  ggtitle("GB Total Wind Speed") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

tws_gom + tws_gb + plot_layout(ncol = 1)
```

#### Helicity ("storminess")

```{r NE-Helicity, echo = F, fig.width= 8, fig.height=8}
gom_helicity <- all_winds %>%
  filter(EPU == "GOM",
         str_detect(Var, "hcly")) %>% #filter
  ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab(expression("Relative Helicity (m"^2*" sec"^-2*")")) +
  ggtitle("GOM Storminess") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

gb_helicity <- all_winds %>%
  filter(EPU == "GB",
         str_detect(Var, "hcly")) %>% #filter
  ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab(expression("Relative Helicity (m"^2*" sec"^-2*")")) +
  ggtitle("GB Storminess") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

gom_helicity + gb_helicity + plot_layout(ncol = 1)
```

#### Turbulent kinetic energy

```{r NE-TKE, echo  = F, fig.width=8, fig.height=8}
gom_tke <- all_winds %>%
  filter(EPU == "GOM",
         str_detect(Var, "tke")) %>% #filter
  ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab(expression("Kinetic Energy (J kg"^-1*")")) +
  ggtitle("GOM Turbulent Kinetic Energy") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

gb_tke <- all_winds %>%
  filter(EPU == "GB",
         str_detect(Var, "tke")) %>% #filter
  ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value))+
  geom_point(aes(x = Time, y = Value))+
  geom_gls(aes(x = Time, y = Value)) +
  ylab(expression("Kinetic Energy (J kg"^-1*")")) +
  ggtitle("GB Turbulent Kinetic Energy") +
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Season ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

gom_tke + gb_tke + plot_layout(ncol = 1)

```

### Ocean Temperature {.tabset .tabset-fade}

#### SST 

```{r NE-SST-insitu, fig.width=8, fig.height = 8}
#EPU shapefile
epu_sf <- ecodata::epu_sf %>% 
  filter(EPU %in% c("GOM","GB"))

#Map line parameters
map.lwd <- 0.4

# Set lat/lon window for maps
xmin = -73
xmax = -65
ymin = 39
ymax = 45
xlims <- c(xmin, xmax)
ylims <- c(ymin, ymax)
sst <- seasonal_oisst_anom_gridded 

sst$Season <- factor(sst$Season, levels = c("Winter",
                                            "Spring",
                                            "Summer",
                                            "Fall"))
sst_map <- 
  ggplot() +
  geom_tile(data = sst, aes(x = Longitude, y = Latitude,fill = Value)) +
  geom_sf(data = coast, size = map.lwd) +
  geom_sf(data = epu_sf, fill = "transparent", size = map.lwd) +
  scale_fill_gradient2(name = "Temp.\nAnomaly (°C)",
                       low = scales::muted("blue"),
                       mid = "white",
                       high = scales::muted("red"),
                       limits = c(-5,5)) +
  coord_sf(crs = crs, xlim = xlims, ylim = ylims) +
  facet_wrap(Season~.) +
  theme_map() +
  ggtitle("SST anomaly (2018)") +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=0.75),
        legend.key = element_blank(),
        axis.title = element_text(size = 11),
        strip.background = element_blank(),
        strip.text=element_text(hjust=0),
        axis.text = element_text(size = 8))


sst_map 

```

```{r NE-SST-anom-series, fig.width = 8}
ne_anom <- seasonal_oisst_anom %>% 
  filter(EPU %in% c("GOM","GB")) %>% 
    mutate(hline = 0,
           Var = str_to_title(str_extract(Var,"winter|spring|summer|fall")))
ne_anom$Var <- factor(ne_anom$Var, levels= c("Winter","Spring","Summer","Fall"))

ne_anom_plt <- ggplot(data = ne_anom, 
       aes(x = Time, y = Value, color = EPU, group = EPU))+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line()+
  geom_point()+
  ylim(-2,3)+
  geom_gls() +
  ylab(expression("SST Anomaly (°C)")) +
  ggtitle("Gulf of Maine & Georges Bank SST Anomaly") +
    scale_color_manual(values = c("black","indianred"))+
    scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  facet_wrap(Var ~., ncol = 2, scales = "free_y")+
  theme_facet() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))
ne_anom_plt
```


#### Bottom temperature

```{r NE-bot-temp}
bot_temp_insitu_gom <- temp_anom %>%
  filter(EPU == "GOM",
         Var == "bottom temp anomaly in situ") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Temperature (°C)") +
  ggtitle("GOM Bot. temp. anomaly (in situ)") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))


bot_temp_insitu_gb <- temp_anom %>%
  filter(EPU == "GB",
         Var == "bottom temp anomaly in situ") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Temperature (°C)") +
  ggtitle("GB Bot. temp. anomaly (in situ)") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))



bot_temp_insitu_gom +
 bot_temp_insitu_gb + 
  plot_layout(ncol =  1) &
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

```

### Ocean Salinity & Stratification {.tabset .tabset-fade}

#### Surface salinity

```{r NW-surfsal}
sal_anom <- oceansal_insitu %>% 
  filter(EPU %in% c("GOM","GB")) %>% 
  complete(Time = full_seq(min(oceansal_insitu$Time):max(oceansal_insitu$Time),1),
           nesting(Var)) %>% 
  mutate(hline = 0)

surf_sal_insitu_gom <- sal_anom %>%
  filter(Var == "surface salinity anomaly in situ",
         EPU == "GOM") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Salinity (PSU)") +
  ggtitle("GOM Surface salinity anomaly (in situ)") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

surf_sal_insitu_gb <- sal_anom %>%
  filter(Var == "surface salinity anomaly in situ",
         EPU == "GB") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Salinity (PSU)") +
  ggtitle("GB Surface salinity anomaly (in situ)") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

sal_oi <- oceansal_oi %>% 
  filter(EPU %in% c("GOM","GB")) %>%
  group_by(EPU, Var) %>% 
  mutate(hline = mean(Value))

surf_sal_oi_GOM <- sal_oi %>%
  filter(Var == "surface sal OI",
         EPU == "GOM") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab(expression("")) +
  ggtitle("GOM Surface salinity (OI)") +
  scale_x_continuous(expand = c(0.01, 0.01), limits = c(1977,2018)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

surf_sal_oi_GB <- sal_oi %>%
  filter(Var == "surface sal OI",
         EPU == "GB") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab(expression("")) +
  ggtitle("GB Surface salinity (OI)") +
  scale_x_continuous(expand = c(0.01, 0.01), limits = c(1977,2018)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))


surf_sal_insitu_gom + surf_sal_oi_GOM +
  surf_sal_insitu_gb + surf_sal_oi_GB +
  plot_layout(ncol =  2) &
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

```

#### Bottom salinity

```{r NE-botsal}

bot_sal_insitu_gom <- sal_anom %>%
  filter(Var == "bottom salinity anomaly in situ",
         EPU == "GOM") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Salinity (PSU)") +
  ggtitle("GOM Bot. salinity anomaly (in situ)") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

bot_sal_insitu_gb <- sal_anom %>%
  filter(Var == "bottom salinity anomaly in situ",
         EPU == "GB") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab("Salinity (PSU)") +
  ggtitle("GB Bot. salinity anomaly (in situ)") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

bot_sal_oi_GOM <- sal_oi %>%
  filter(Var == "bottom sal OI",
         EPU == "GOM") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab(expression("")) +
  ggtitle("GOM Bot. salinity (OI)") +
  scale_x_continuous(expand = c(0.01, 0.01), limits = c(1977,2018)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))

bot_sal_oi_GB <- sal_oi %>%
  filter(Var == "bottom sal OI",
         EPU == "GB") %>% 
    ggplot()+ #plot
     annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line(aes(x = Time, y = Value)) +
  geom_gls(aes(x = Time, y = Value)) +
  geom_point(aes(x = Time, y = Value), size = 1) +
  ylab(expression("")) +
  ggtitle("GB Bot. salinity (OI)") +
  scale_x_continuous(expand = c(0.01, 0.01), limits = c(1977,2018)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() +
  theme(strip.text=element_text(hjust=0),
        plot.title = element_text(size = 12))


bot_sal_insitu_gom + bot_sal_oi_GOM +
  bot_sal_insitu_gb + bot_sal_oi_GB +
  plot_layout(ncol =  2) &
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

```

#### Stratification

```{r NE-strat, fig.width = 8, fig.asp = .35}
strat <- ecodata::stratification %>% 
  filter(EPU %in% c("GOM","GB")) %>% 
  group_by(EPU) %>% 
  mutate(hline = mean(Value, na.rm = T))

gom_strat <- strat %>% 
  filter(EPU == "GOM") %>% 
  ggplot(aes(x = Time, y = Value)) +
       annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line() +
  geom_point() +
  geom_gls() +
    ylab(expression("Stratification (kg m"^-3*")")) +
  ggtitle("GOM Stratification") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() 

gb_strat <- strat %>% 
  filter(EPU == "GB") %>% 
  ggplot(aes(x = Time, y = Value)) +
       annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line() +
  geom_point() +
  geom_gls() +
    ylab(expression("Stratification (kg m"^-3*")")) +
  ggtitle("GB Stratification") +
  scale_x_continuous(expand = c(0.01, 0.01)) +
    geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts() 

gom_strat + gb_strat + plot_layout(ncol = 2)
```

#### GOM Slope Water Proportions

```{r wsw-prop, fig.width=5, fig.asp = 0.55}

sw.df <- slopewater %>% 
  mutate(Var, Var = plyr::mapvalues(Var, from = c("WSW proportion ne channel",
                                                  "LSLW proportion ne channel"),
                                    to = c("WSW","LSLW"))) %>% 
  dplyr::rename(Flavor  = Var) %>% 
  group_by(Flavor) %>% 
  mutate(hline = mean(Value)) 

sw.df$Flavor <- factor(sw.df$Flavor, levels = c("WSW","LSLW"))

ggplot(data = sw.df) +
  geom_line(aes(x = Time, y = Value, color = Flavor))+
  geom_point(aes(x = Time, y = Value, color = Flavor)) +
  ylab("Percent of Total Slopewater") +
  ggtitle("Slopewater Proportions in NE Channel")+
    scale_x_continuous(expand = c(0.01, 0.01))+
      geom_hline(aes(yintercept = hline,
                     color = Flavor),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_ts() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))

```


### Chlorophyll and Primary Productivity {.tabset .tabset-fade}

#### Primary production

```{r NE-PP-OCCI, fig.width=8, fig.height = 8}
out_pp <- ecodata::chl_pp %>% 
  filter(EPU %in% c("GOM","GB"),
         str_detect(Var, "MONTHLY_PPD_MEDIAN")) %>% 
  separate(.,Time, into = c("Year","Month"), sep = 4) %>% 
    mutate(Month = plyr::mapvalues(Month, from = c("01","02","03","04","05","06",
                                                   "07","08","09","10","11","12"),
                                   to = c(month.abb))) %>% 
  group_by(EPU, Month) %>% 
  mutate(hline = mean(Value))
out_pp$Month <- factor(out_pp$Month, levels = month.abb)

pp_cci_gom <- out_pp %>% 
  filter(EPU == "GOM") %>% 
 ggplot() +
   # geom_gls(aes(x = Year, y = Value, group = Month))+
    geom_point(aes(x = Year, y = Value, group = Month)) +
    geom_line(aes(x = Year, y = Value, group = Month)) +
    scale_x_discrete(name = "Time", breaks = seq(min(out_pp$Year),max(out_pp$Year),10)) +  
    facet_wrap(Month~., ncol = 6) +
    ggtitle("GOM Monthly median PPD") +
    ylab(expression("PP (gC m"^-2*" d"^-1*")")) +
        geom_hline(aes(yintercept = hline,
                     group = Month),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
    theme_facet() +
    theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm"))
 
 pp_cci_gb <-out_pp %>% 
  filter(EPU == "GB") %>% 
 ggplot() +
   # geom_gls(aes(x = Year, y = Value, group = Month))+
    geom_point(aes(x = Year, y = Value, group = Month)) +
    geom_line(aes(x = Year, y = Value, group = Month)) +
    scale_x_discrete(name = "Time", breaks = seq(min(out_pp$Year),max(out_pp$Year),10)) +  
    facet_wrap(Month~., ncol = 6) +
    ggtitle("GB Monthly median PPD") +
    ylab(expression("PP (gC m"^-2*" d"^-1*")")) +
        geom_hline(aes(yintercept = hline,
                     group = Month),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
    theme_facet() +
    theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm"))
 
 pp_cci_gom + pp_cci_gb + plot_layout(ncol = 1)
 
```

#### Sesonal chlorophyll *a* & primary production

```{r NE-chl-weekly, fig.width = 8, fig.cap = "Weekly chlorophyll concentrations and primary productivity for 2018 in Gulf of Maine and Georges Bank are shown by the colored lines in the above figures. The long-term mean is shown in black and shading indicates +/- 1 sample SD."}

GB_chl_weekly <- interp_chl_pp(epu = "GB", int_step = 1000, year = 2018,Variable = "WEEKLY_CHLOR_A_MEDIAN")

GB_chl <- ggplot(data = GB_chl_weekly) +
  geom_line(aes(x = Time, y = LTM)) +
  geom_ribbon(aes(x = Time, ymin = pmax(sd.low,0), ymax = sd.high), 
              alpha = 0.1,
              fill = "grey1") +
  geom_line(aes(x = Time, y = Value, color = status, group = group),
            size = 1) +
  ggtitle(expression("GB chlorophyll"~italic(a)~"")) +
  guides(color = F) +
  xlab("")+
  ylab("") +
  scale_x_continuous(breaks = seq(0,50,10),
                   labels = c("January","March","May","July","October","December")) +
  scale_color_manual(values = c("#ef8a62","#2c7fb8","#a1d99b"))+
  theme_ts()


GOM_chl_weekly <- interp_chl_pp(epu = "GOM", int_step = 1000, year = 2018,Variable = "WEEKLY_CHLOR_A_MEDIAN")

GOM_chl <- ggplot(data = GOM_chl_weekly) +
  geom_line(aes(x = Time, y = LTM)) +
  geom_ribbon(aes(x = Time, ymin = sd.low, ymax = sd.high), 
              alpha = 0.1,
              fill = "grey1") +
  geom_line(aes(x = Time, y = Value, color = status, group = group),
            size = 1) +
  ggtitle(expression("GOM chlorophyll"~italic(a)~"")) +
  guides(color = F) +
  xlab("")+
  ylab(expression("CHL (mg m"^-3*")")) +
  scale_x_continuous(breaks = seq(0,50,10),
                   labels = c("January","March","May","July","October","December")) +
  scale_color_manual(values = c("#ef8a62","#2c7fb8","#a1d99b"))+
  theme_ts()

GB_ppd_weekly <- interp_chl_pp(epu = "GB", int_step = 1000, year = 2018,Variable = "WEEKLY_PPD_MEDIAN")

GB_ppd <- ggplot(data = GB_ppd_weekly) +
  geom_line(aes(x = Time, y = LTM)) +
  geom_ribbon(aes(x = Time, ymin = pmax(sd.low,0), ymax = sd.high), 
              alpha = 0.1,
              fill = "grey1") +
  geom_line(aes(x = Time, y = Value, color = status, group = group),
            size = 1) +
  ggtitle(expression("GB primary production")) +
  guides(color = F) +
  xlab("")+
  ylab("") +
  scale_x_continuous(breaks = seq(0,50,10),
                   labels = c("January","March","May","July","October","December")) +
  scale_color_manual(values = c("#ef8a62","#2c7fb8","#a1d99b"))+
  theme_ts()


GOM_ppd_weekly <- interp_chl_pp(epu = "GOM", int_step = 1000, year = 2018,Variable = "WEEKLY_PPD_MEDIAN")

GOM_ppd <- ggplot(data = GOM_ppd_weekly) +
  geom_line(aes(x = Time, y = LTM)) +
  geom_ribbon(aes(x = Time, ymin = sd.low, ymax = sd.high), 
              alpha = 0.1,
              fill = "grey1") +
  geom_line(aes(x = Time, y = Value, color = status, group = group),
            size = 1) +
  guides(color = F) +
  xlab("")+
  ggtitle(expression("GOM primary production")) +
  scale_x_continuous(breaks = seq(0,50,10),
                   labels = c("January","March","May","July","October","December")) +
  ylab((expression("PP (gC m"^-2*" d"^-1*")"))) +
  scale_color_manual(values = c("#ef8a62","#2c7fb8","#a1d99b"))+
  theme_ts()


GOM_chl + GB_chl + GOM_ppd + GB_ppd + plot_layout(ncol = 2) & theme(plot.margin = margin(r = 0.5))

```



```{r NE-chl-monthly, fig.width = 8, fig.height=8, eval = F}
out_chl <- ecodata::chl_pp %>% 
  filter(EPU %in% c("GOM","GB"),
         str_detect(Var, "MONTHLY_CHLOR_A_MEDIAN")) %>% 
  separate(.,Time, into = c("Year","Month"), sep = 4) %>% 
    mutate(Month = plyr::mapvalues(Month, from = c("01","02","03","04","05","06",
                                                   "07","08","09","10","11","12"),
                                   to = c(month.abb)))
out_chl$Month <- factor(out_chl$Month, levels = month.abb)


chl_cci_gom <- 
  out_chl %>% 
  filter(EPU == "GOM") %>% 
  ggplot() +
    geom_gls(aes(x = Year, y = Value, group = Month))+
    geom_point(aes(x = Year, y = Value, group = Month)) +
    geom_line(aes(x = Year, y = Value, group = Month)) +
    scale_x_discrete(name = "Time", breaks = seq(min(out_chl$Year),max(out_chl$Year),10)) +  
    facet_wrap(Month~., ncol = 6) +
    ggtitle("GOM Monthly median CHL (OC-CCI)") +
    ylab(expression("CHL (mg m"^-3*")")) +
    theme_facet() +
    theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm"))

chl_cci_gb <- 
  out_chl %>% 
  filter(EPU == "GB") %>% 
  ggplot() +
    geom_gls(aes(x = Year, y = Value, group = Month))+
    geom_point(aes(x = Year, y = Value, group = Month)) +
    geom_line(aes(x = Year, y = Value, group = Month)) +
    scale_x_discrete(name = "Time", breaks = seq(min(out_chl$Year),max(out_chl$Year),10)) +  
    facet_wrap(Month~., ncol = 6) +
    ggtitle("GB Monthly median CHL (OC-CCI)") +
    ylab(expression("CHL (mg m"^-3*")")) +
    theme_facet() +
    theme(axis.text.x = element_text(angle=45, hjust = 1),
          panel.spacing = unit(1, "lines"),
          plot.margin = unit(c(0.1, 0, 0, 0), "cm"))
chl_cci_gom + chl_cci_gb + plot_layout(ncol = 1)

```

### Zooplankton {.tabset .tabset-fade}

#### Abundance anomaly

```{r NE-zoo-abund}
zoo_abund <- ecodata::zoo_anom_sli %>% 
  filter(EPU %in% c("GOM","GB"),
         !str_detect(Var, "small-large")) %>% 
  mutate(hline = 0,
         Var = str_to_title(str_remove(Var, "anomaly"))) 

gom_zoo_abund <- zoo_abund %>% 
  filter(EPU == "GOM") %>% 
  ggplot(aes(x = Time, y = Value, group = Var)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_gls() +
  geom_line() +
  geom_point() +
  ylab("Abundance anomaly") +
  ggtitle("GOM Zooplankton abundance anomaly") +
  facet_wrap(Var~., ncol = 3) +
  scale_x_continuous(expand = c(0.01, 0.01))+
      geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_facet() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))

gb_zoo_abund <- zoo_abund %>% 
  filter(EPU == "GB") %>% 
  ggplot(aes(x = Time, y = Value, group = Var)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_gls() +
  geom_line() +
  geom_point() +
  ylab("Abundance anomaly") +
  ggtitle("GB Zooplankton abundance anomaly") +
  facet_wrap(Var~., ncol = 3) +
  scale_x_continuous(expand = c(0.01, 0.01))+
      geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_facet() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))
  
gom_zoo_abund + gb_zoo_abund + plot_layout(ncol = 1)

```

#### Small-large index

```{r NE-sli}
sli <- ecodata::zoo_anom_sli %>% 
  filter(EPU %in% c("GOM","GB"),
         str_detect(Var, "small-large")) %>% 
  mutate(hline = 0) 
sli$EPU <- factor(sli$EPU, levels = c("GOM","GB"))

pp_anom <- ecodata::chl_pp %>% 
  filter(str_detect(Var, "ANNUAL_PPD_RATIO_ANOMALY"),
         EPU %in% c("GOM","GB")) %>% 
  mutate(hline = 1,
         Time = as.numeric(as.character(Time)),
         Var = "ANNUAL_PPD_RATIO_ANOMALY")
pp_anom$EPU <- factor(pp_anom$EPU, levels = c("GOM","GB"))

sli_plt <- sli %>% 
  ggplot(aes(x = Time, y = Value, group = Var)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line() +
  geom_point() +
  guides(color = F) +
  facet_wrap(EPU~.,ncol = 2)+
  xlab("")+
  ylab("Small-large abundance") +
  ggtitle("Small-large copepod abundance") +
    scale_x_continuous(expand = c(0.01, 0.01), limits = c(1998, 2018))+
      geom_hline(aes(yintercept = hline,
                     group = Var),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_facet() +
  theme(strip.text=element_text(hjust=0))

pp_anom_plt <- pp_anom %>% 
    ggplot(aes(x = Time, y = Value, group = Var)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_line() +
  geom_point() +
  guides(color = F) +
  ylab("Anomaly ratio") +
  facet_wrap(EPU~.,ncol = 2)+
  ggtitle("Primary production anomaly ratio") +
    scale_x_continuous(expand = c(0.01, 0.01), limits = c(1998, 2018))+
    scale_y_continuous(limits = c(0.8,1.2)) +
      geom_hline(aes(yintercept = hline,
                     group = Var),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_facet() +
  theme(strip.text=element_text(hjust=0))


sli_plt + pp_anom_plt + plot_layout(ncol = 1) & theme(plot.margin = margin(t = 0))
```

#### OI abundance

```{r NE-oi-zoo, fig.width=8, fig.height=9}
facet_names <- list(
  'centropages spring'=expression(paste(italic("Centropages "), "spring")),
  'centropages fall'=expression(paste(italic("Centropages "), "fall")),
  'temora spring'=expression(paste(italic("Temora "), "spring")),
  'temora fall' = expression(paste(italic("Temora "), "fall")),
  'pseudocalanus spring'=expression(paste(italic("Pseudocalanus "), "spring")),
  'pseudocalanus fall' = expression(paste(italic("Pseudocalanus "), "fall")))

zoo_oi_ne <- ecodata::zoo_oi %>% 
  filter(EPU %in% c("GOM","GB"),
         !str_detect(Var,"SD")) %>% 
  mutate(Var = str_remove(Var, " zoo")) %>% 
  group_by(EPU, Var) %>% 
  mutate(hline = mean(Value, na.rm = T))

gom_zoo <- zoo_oi_ne %>% 
  filter(EPU == "GOM") %>% 
ggplot(aes(x = Time, y = Value, group = Var)) +
  geom_gls() +
  geom_line()+
  geom_point()+
  ylab(expression("Abundance log num m"^-3*"")) +
  facet_wrap(Var ~ ., ncol = 2,scales = "free_y", labeller = label) +
  ggtitle("GOM Zooplankton abundance (OI)") +
  theme_facet() +
    theme(strip.text=element_text(hjust=0,
                                face = "italic"))

gb_zoo <- zoo_oi_ne %>% 
  filter(EPU == "GB") %>% 
ggplot(aes(x = Time, y = Value, group = Var)) +
  geom_gls() +
  geom_line()+
  geom_point()+
  ylab(expression("Abundance log num m"^-3*"")) +
  facet_wrap(Var ~ ., ncol = 2,scales = "free_y", labeller = label) +
  ggtitle("GB Zooplankton abundance (OI)") +
  theme_facet() +
    theme(strip.text=element_text(hjust=0,
                                face = "italic"))
gom_zoo + gb_zoo + plot_layout(ncol = 1)

```


## Shelfwide Indicators {.tabset .tabset-fade}

### Seasonal SST anomaly

```{r seasonal-sst, fig.height=8}
# Set lat/lon window for maps
xmin = -77
xmax = -65
ymin = 36
ymax = 45
xlims <- c(xmin, xmax)
ylims <- c(ymin, ymax)
sst <- seasonal_oisst_anom_gridded 

sst$Season <- factor(sst$Season, levels = c("Winter",
                                            "Spring",
                                            "Summer",
                                            "Fall"))

ggplot() +
  geom_tile(data = sst, aes(x = Longitude, y = Latitude,                                             fill = Value)) +
  geom_sf(data = coast, size = map.lwd) +
  geom_sf(data = epu_sf, fill = "transparent", size = map.lwd) +
  scale_fill_gradient2(name = "Temp.\nAnomaly (°C)",
                       low = scales::muted("blue"),
                       mid = "white",
                       high = scales::muted("red"),
                       limits = c(-5,5)) +
  coord_sf(crs = crs, xlim = xlims, ylim = ylims) +
  facet_wrap(Season~.) +
  theme_map() +
  ggtitle("SST anomaly (2018)") +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=0.75),
        legend.key = element_blank(),
        axis.title = element_text(size = 11),
        strip.background = element_blank(),
        strip.text=element_text(hjust=0),
        axis.text = element_text(size = 8))
```

### Long-term SST

```{r long-term-sst}
lt_sst <- ecodata::long_term_sst %>% 
  mutate(hline = mean(Value))

lt_sst %>% 
  ggplot(aes(x = Time, y = Value, group = Var)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_gls() +
  geom_line() +
  geom_point() +
  ylab("Temperature (°C)") +
  ggtitle("Long-term SST") +
    scale_x_continuous(expand = c(0.01, 0.01), breaks = seq(1840,2010,10))+
      geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_facet() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))
```

### NAO

```{r NAO, fig.cap = "Wintertime (Dec-Mar) differences of normalized sea level pressures between Lisbon and Stykkisholmur/Reykjavik, Iceland since 1864."}

# [https://climatedataguide.ucar.edu/climate-data/hurrell-north-atlantic-oscillation-nao-index-station-based](source)
# higher pressure in iceland, more negative index
nao %>% 
  mutate(hline = 0,
         status = ifelse(Value < 0, "negative","positive")) %>% 
  ggplot(aes(x = Time, y = Value, fill = status)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_bar(stat = "identity") +
  guides(fill = F)+
  ylab("NAO Index") +
  ggtitle("NAO Index") +
    scale_x_continuous(expand = c(0.01, 0.01), breaks = seq(1840,2010,10))+
      geom_hline(aes(yintercept = hline),
           alpha = hline.alpha)+
  ylab(expression(Lisbon["n"]~-~Stykkisholmur/Reykjavik["n"]))+
  theme_ts() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))

# nao_var <- nao_ann %>% 
#   ggplot(aes(x = Time, y = Variance)) +
#          annotate("rect", fill = shade.fill, alpha = shade.alpha,
#       xmin = x.shade.min , xmax = x.shade.max,
#       ymin = -Inf, ymax = Inf) +
#   geom_gls() +
#   geom_line() +
#   geom_point() +
#   ylab("Annual Variance") +
#   ggtitle("NAO Index Variance") +
#     scale_x_continuous(expand = c(0.01, 0.01))+
#       geom_hline(aes(yintercept = hline.var),
#            size = hline.size,
#            alpha = hline.alpha,
#            linetype = hline.lty)+
#   theme_ts() +
#   theme(strip.text=element_text(hjust=0,
#                                 face = "italic"))
# 
# nao_in + nao_var + plot_layout(ncol = 2) &
#   theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
```

### GSI

```{r GSI, fig.width = 5, fig.asp = 0.55}
gsi %>% 
  mutate(Year = floor(Time)) %>% 
  group_by(Year) %>% 
  dplyr::summarise(Value = mean(Value)) %>% 
  mutate(hline = mean(Value)) %>% 
  dplyr::rename(Time = Year) %>% 
  ggplot(aes(x = Time, y = Value)) +
         annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  geom_gls() +
  geom_line() +
  geom_point() +
  ylab("Gulf Stream position anomaly") +
  ggtitle("Gulf Stream Index") +
    scale_x_continuous(expand = c(0.01, 0.01))+
      geom_hline(aes(yintercept = hline),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty)+
  theme_ts() +
  theme(strip.text=element_text(hjust=0,
                                face = "italic"))
```

