# Regime Shifts case studies 

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center') #allows for inserting R code into captions

#Plotting and data libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(ecodata)

#Time series constants
shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
feeding.guilds <- c("Apex Predator","Piscivore","Planktivore","Benthivore","Benthos")
x.shade.min <- 2011
x.shade.max <- 2021
series.col <- c("black","indianred")
#Function for custom ggplot facet labels
label <- function(variable,value){
  return(facet_names[value])
}

```

## Data included {.tabset .tabset-fade}

Table describing the data included in the regime shift case studies. 

| EPU | Indicator                          | Year_min | ecodata::           |
|-----|------------------------------------|----------|---------------------|
| MAB | Fall SST                           | 1982     | seasonal_oisst_anom |
| MAB | Zooplankton                        | 1977     | zoo_regime           |   
| MAB | Forage Index                       | 1985     |  forage_index      |
| MAB | Total Productivity                 | 1980     | productivity_anomaly|
| GOM | Annual Surface Cumulative Heatwave | 1982     | heatwave            |
| GOM | Zooplankton                        | 1977     | zoo_regime            |
| GOM | Forage Index                       | 1985     | forage_index       |
| GOM | Condition Index                    | 1992     | condition          |


### Mid-Atlantic

```{r}
temp <- ecodata::seasonal_oisst_anom %>%
  dplyr::filter(EPU == "MAB", 
                Var == "Fall") %>% 
  dplyr::rename("anom" = "Value") %>% 
  #dplyr::select(!Units) %>% 
  dplyr::mutate(Var = recode(Var, "Fall" ="Temperature"))

zoo<- ecodata::zoo_regime %>% 
  dplyr::filter(EPU == "MAB") %>% 
  dplyr::group_by(Time) %>% 
  dplyr::summarise(anom = mean(Value)) %>% 
  dplyr::mutate(Var = "Zooplankton", 
                EPU = "MAB")

forage<- ecodata::forage_index %>% 
         dplyr::filter(EPU == "MAB",
                       Var == "Spring Forage Fish Biomass Estimate") %>%
         #dplyr::rename("estimate" = 'Fall Forage Fish Biomass Estimate') %>% 
         dplyr::mutate(anom = Value - mean(Value)) %>% 
  dplyr::select(!c(Value,Units))%>% 
  dplyr::mutate(Var = recode(Var, "Spring Forage Fish Biomass Estimate" ="Forage"))


productivity<- ecodata::productivity_anomaly %>% 
  dplyr::filter(EPU == "MAB") %>% 
  dplyr::group_by(Time) %>%
  dplyr::summarise(anom = sum(Value, na.rm = T)) %>% 
  dplyr::mutate(anom = ifelse(anom == 0, NA, anom), 
                Var = "Productivity", 
                EPU = "MAB")

mab <- rbind(temp, zoo, forage, productivity)
mab$Var <- factor(mab$Var,levels = c("Temperature",
                                     "Zooplankton", 
                                     "Forage", 
                                     "Productivity"))

mab %>% 
  ggplot2::ggplot(aes(x = Time, y = anom)) +
  ggplot2::annotate("rect", fill = shade.fill, alpha = shade.alpha,
                    xmin = x.shade.min , xmax = x.shade.max,
                    ymin = -Inf, ymax = Inf) +
  ecodata::geom_regime()+
  ggplot2::geom_col(aes(fill = ifelse(anom<0, "red", "blue"))) +
  ggplot2::ylab("Anomaly") +
  ggplot2::xlab(element_blank())+
  facet_wrap(.~Var, ncol = 1, scales = "free_y")+
  #ggplot2::ggtitle("Productivity anomaly") +
  ggplot2::scale_x_continuous(expand = c(0.01, 0.01))+
  ecodata::theme_ts() +
  ggplot2::theme(strip.text=element_text(hjust=0,
                                         face = "italic"), 
                 legend.position = "none")

```


#### Temperature

```{r}
ecodata::seasonal_oisst_anom %>%
  dplyr::filter(EPU == "MAB", 
                Var == "Fall") %>% 
  ggplot2::ggplot(aes(x = Time, y = Value)) +
  ggplot2::annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  ecodata::geom_regime()+
  ggplot2::geom_col(aes(fill = ifelse(Value<0, "red", "blue"))) +
  ggplot2::ylab("Anomaly") +
  ggplot2::xlab(element_blank())+
  ggplot2::ggtitle("Fall") +
  ggplot2::scale_x_continuous(expand = c(0.01, 0.01))+
  ecodata::theme_ts() +
  ggplot2::theme(strip.text=element_text(hjust=0,
                                face = "italic"), 
                 legend.position = "none")
```


#### Zooplankton 

```{r}
reg<- ecodata::zoo_regime %>% 
  dplyr::filter(EPU == "MAB") %>% 
  dplyr::group_by(Time) %>% 
  dplyr::mutate(hline = mean(Value))

ecodata::zoo_regime %>% 
  dplyr::filter(EPU == "MAB") %>% 
   ggplot(aes(x = Time, y = Value, fill = Var)) +
  ggtitle("Zooplankton Abundance anomaly") + 
  geom_bar(stat='identity') +
  ggplot2::annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  ecodata::geom_regime(data = reg,aes(x = Time, y = Value) )

```

#### Forage Index

```{r}
forage<- ecodata::forage_index %>% 
  dplyr::filter(EPU == "MAB", 
                stringr::str_detect(Var,"Fall")) %>% 
  tidyr::pivot_wider(names_from = "Var", values_from = "Value") %>% 
  dplyr::rename("estimate" = 'Fall Forage Fish Biomass Estimate', 
                "se" = 'Fall Forage Fish Biomass Estimate SE') %>% 
  dplyr::mutate(upper = estimate + se, 
                lower = estimate - se) %>% 
  ggplot2::ggplot(aes(x = Time, y = estimate))+
  ggplot2::geom_point()+
  ggplot2::geom_line()+
  ggplot2::geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  ecodata::geom_regime()+
  ecodata::theme_ts()+
  ggtitle("Fall Forage Index - Mid- Atlantic")
  
  
forage

```

```{r}
forage<- ecodata::forage_index %>% 
  dplyr::filter(EPU == "MAB",
                Var == "Fall Forage Fish Biomass Estimate") %>%
    #dplyr::rename("estimate" = 'Fall Forage Fish Biomass Estimate') %>% 
  dplyr::mutate(anom = Value - mean(Value)) %>% 
  ggplot2::ggplot(aes(x = Time, y = anom))+
  #ggplot2::geom_point()+
  #ggplot2::geom_line()+
  ggplot2::geom_col(aes(fill = ifelse(anom<0, "red", "blue"))) +
  #ggplot2::geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  ecodata::geom_regime()+
  ecodata::theme_ts()+
  ggtitle("Fall Forage Index - Mid- Atlantic")+
  ggplot2::theme(strip.text=element_text(hjust=0,
                                face = "italic"), 
                 legend.position = "none")
  
forage

```



#### Productivity

```{r}
mid <- ecodata::productivity_anomaly %>% 
  dplyr::filter(EPU == "MAB") %>% 
  dplyr::group_by(Time) %>%
     dplyr::summarise(Total = sum(Value, na.rm = T)) %>% 
     dplyr::mutate(Total = ifelse(Total == 0, NA, Total))

mid %>% ggplot2::ggplot(aes(x = Time, y = Total)) +
  ggplot2::annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  ecodata::geom_regime()+
  ggplot2::geom_col(aes(fill = ifelse(Total<0, "red", "blue"))) +
  ggplot2::ylab("Anomaly") +
  ggplot2::xlab(element_blank())+
  ggplot2::ggtitle("Productivity anomaly") +
  ggplot2::scale_x_continuous(expand = c(0.01, 0.01))+
  ecodata::theme_ts() +
  ggplot2::theme(strip.text=element_text(hjust=0,
                                face = "italic"), 
                 legend.position = "none")
```


### Gulf of Maine

```{r}
temp <- ecodata::seasonal_oisst_anom %>%
  dplyr::filter(EPU == "GOM", 
                Var == "Fall") %>% 
  dplyr::rename("anom" = "Value") %>% 
  #dplyr::select(!Units) %>% 
  dplyr::mutate(Var = recode(Var, "Fall" ="Temperature"))


hw <- ecodata::heatwave %>%
  dplyr::filter(EPU == "GOM", 
                Var == "cumulative intensity-Surface") %>% 
  dplyr::mutate(anom = Value/(mean(Value))-1) %>% 
  dplyr::select(!c(Units, Value)) %>% 
  dplyr::mutate(Var = recode(Var, "cumulative intensity-Surface" ="Heatwave"))


zoo<- ecodata::zoo_regime %>% 
  dplyr::filter(EPU == "GOM") %>% 
  dplyr::group_by(Time) %>% 
  dplyr::summarise(anom = mean(Value)) %>% 
  dplyr::mutate(Var = "Zooplankton", 
                EPU = "GOM")

forage<- ecodata::forage_index %>% 
         dplyr::filter(EPU == "GOM",
                       Var == "Spring Forage Fish Biomass Estimate") %>%
         #dplyr::rename("estimate" = 'Fall Forage Fish Biomass Estimate') %>% 
         dplyr::mutate(anom = Value - mean(Value)) %>% 
  dplyr::select(!c(Value,Units))%>% 
  dplyr::mutate(Var = recode(Var, "Spring Forage Fish Biomass Estimate" ="Forage"))

condition<- ecodata::condition %>% 
  ungroup() %>% 
  dplyr::group_by(Time) %>% 
  dplyr::summarise(anom = mean(Value) - 1)%>% 
  dplyr::mutate(Var = "Condition", 
                EPU = "GOM")

gom <- rbind(temp, hw, zoo, forage, condition) 

gom$Var <- factor(gom$Var,levels = c("Temperature",
                                     "Heatwave",  
                                     "Zooplankton", 
                                     "Forage", 
                                     "Condition"))
gom %>%   ggplot2::ggplot(aes(x = Time, y = anom)) +
  ggplot2::annotate("rect", fill = shade.fill, alpha = shade.alpha,
                    xmin = x.shade.min , xmax = x.shade.max,
                    ymin = -Inf, ymax = Inf) +
  ecodata::geom_regime()+
  ggplot2::geom_col(aes(fill = ifelse(anom<0, "red", "blue"))) +
  ggplot2::ylab("Anomaly") +
  ggplot2::xlab(element_blank())+
  facet_wrap(.~Var, ncol = 1, scales = "free_y")+
  #ggplot2::ggtitle("Productivity anomaly") +
  ggplot2::scale_x_continuous(expand = c(0.01, 0.01))+
  ecodata::theme_ts() +
  ggplot2::theme(strip.text=element_text(hjust=0,
                                         face = "italic"), 
                 legend.position = "none")

```




#### Heatwave

```{r}
ecodata::heatwave %>%
  dplyr::filter(EPU == "GOM", 
                Var == "cumulative intensity") %>% 
  dplyr::mutate(anom = Value/(mean(Value))-1) %>% 
  ggplot2::ggplot(aes(x = Time, y = anom)) +
  ggplot2::annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  ecodata::geom_regime()+
  ggplot2::geom_col(aes(fill = ifelse(anom<0, "red", "blue"))) +
  ggplot2::ylab("Anomaly") +
  ggplot2::xlab(element_blank())+
  ggplot2::ggtitle("Cumulative Heatwave Anomaly") +
  ggplot2::scale_x_continuous(expand = c(0.01, 0.01))+
  ecodata::theme_ts() +
  ggplot2::theme(strip.text=element_text(hjust=0,
                                face = "italic"), 
                 legend.position = "none")
```


#### Zooplankton 



```{r}
reg<- ecodata::zoo_regime %>% 
  dplyr::filter(EPU == "GOM") %>% 
  dplyr::group_by(Time) %>% 
  dplyr::mutate(hline = mean(Value))

ecodata::zoo_regime %>% 
  dplyr::filter(EPU == "GOM") %>% 
   ggplot(aes(x = Time, y = Value, fill = Var)) +
  ggtitle("Zooplankton Abundance anomaly") + 
  geom_bar(stat='identity') +
  ggplot2::annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf) +
  ecodata::geom_regime(data = reg,aes(x = Time, y = hline) )


```

#### Forage Index


```{r}
forage<- ecodata::forage_index %>% 
  dplyr::filter(EPU == "GOM", 
                stringr::str_detect(Var,"Fall")) %>% 
  tidyr::pivot_wider(names_from = "Var", values_from = "Value") %>% 
  dplyr::rename("estimate" = 'Fall Forage Fish Biomass Estimate', 
                "se" = 'Fall Forage Fish Biomass Estimate SE') %>% 
  dplyr::mutate(upper = estimate + se, 
                lower = estimate - se) %>% 
  ggplot2::ggplot(aes(x = Time, y = estimate))+
  ggplot2::geom_point()+
  ggplot2::geom_line()+
  ggplot2::geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  ecodata::geom_regime()+
  ecodata::theme_ts()+
  ggtitle("Fall Forage Index - Gulf of Maine")
  
forage

```

```{r}
forage<- ecodata::forage_index %>% 
  dplyr::filter(EPU == "GOM",
                Var == "Fall Forage Fish Biomass Estimate") %>%
    #dplyr::rename("estimate" = 'Fall Forage Fish Biomass Estimate') %>% 
  dplyr::mutate(anom = Value - mean(Value)) %>% 
  ggplot2::ggplot(aes(x = Time, y = anom))+
  #ggplot2::geom_point()+
  #ggplot2::geom_line()+
  ggplot2::geom_col(aes(fill = ifelse(anom<0, "red", "blue"))) +
  #ggplot2::geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  ecodata::geom_regime()+
  ecodata::theme_ts()+
  ggtitle("Fall Forage Index - Gulf of Maine")+
  ggplot2::theme(strip.text=element_text(hjust=0,
                                face = "italic"), 
                 legend.position = "none")

forage
```

#### Condition

```{r}
cond<- ecodata::condition %>% 
  mutate(anom = Value/(mean(Value))-1) %>% 
  ggplot2::ggplot(aes(x = Time, y = anom))+
  ggplot2::geom_col(aes(fill = ifelse(anom<0, "red", "blue")))+
  ggplot2::facet_wrap(.~Var)+
  ecodata::geom_regime()+
  ggplot2::theme(legend.position = "none")
cond
```


```{r}


cond1<- ecodata::condition %>% 
  # dplyr::filter(Var %in% c("Atlantic cod",
  #                        "Haddock",
  #                        "Yellowtail flounder",
  #                        "Pollock",
  #                        "American plaice",
  #                        "Witch flounder",
  #                        "White hake",
  #                        "Windowpane",
  #                        "Winter flounder",
  #                        "Acadian redfish",
  #                        "Atlantic wolffish",
  #                        "Ocean pout" )) %>%
  mutate(anom = Value/(mean(Value))-1) %>%
  ggplot2::ggplot(aes(x = Time, y = mean))+
  ggplot2::geom_bar(aes(fill = Var, x = Time, y = anom),
                    stat = "identity")+
  ggplot2::theme(legend.position = "none")
cond1


```

```{r}
condup<- ecodata::condition %>% 
  dplyr::mutate(anom = Value/(mean(Value))-1) %>% 
  dplyr::filter(anom > 0) %>% 
  group_by(Time) %>% 
  dplyr::mutate(numb = length(unique(Var)), 
                category = c("above")) %>% 
  ungroup()

ecodata::condition %>% 
  dplyr::mutate(anom = Value/(mean(Value))-1) %>% 
  dplyr::filter(anom < 0) %>% 
  group_by(Time) %>% 
  dplyr::mutate(numb = length(unique(Var)), 
                category = c("below")) %>% 
  ungroup() %>%
  rbind(condup) %>% 
  ggplot2::ggplot(aes(x = Time, y = numb, fill = category ))+
  # geom_point()+
  # geom_line()+
  # geom_point(data = condup, aes(x = Time, y = numb))+
  # geom_line(data = condup, aes(x = Time, y = numb), color = "blue")+
  geom_bar(position="fill", stat="identity")+
  ylab("number of species")+
  ggtitle("Condition Anomaly Species Count")+
  theme(legend.title = element_blank())#+
  #ecodata::geom_regime()

```

```{r}

ecodata::condition %>% 
  dplyr::mutate(anom = Value/(mean(Value))-1) %>% 
  dplyr::filter(anom > 0) %>% 
  group_by(Time) %>% 
  dplyr::summarise(numb = length(unique(Var)), 
                category = c("above")) %>% 
  ungroup() %>% 
  dplyr::mutate(anom = numb - mean(numb), 
                hline = mean(numb)) %>% 
  ggplot2::ggplot(aes(x = Time, y = numb))+
  #geom_point()+
  #geom_line()+
  ecodata::geom_regime()+
  ggplot2::geom_col(aes(fill = ifelse(anom<0, "red", "blue")))+ 
  ylab("# of species")+
  ggplot2::geom_hline(aes(yintercept = unique(hline)))+
  ggtitle("Condition Anomaly Species Count - Number above 0")+
  theme(legend.title = element_blank(), 
        legend.position = "none")



```










